\input{preamble}

% Override title for the short letter
\makeatletter
\renewcommand{\@title}{On-Chain Parametric Cost Insurance for Underserved
Latin-American Digital Workers \\[0.3em]
\large Record of Closed Iterations and Request for Methodological Guidance}
\renewcommand{\@author}{[author block --- fill at composition time]}
\renewcommand{\@date}{2026-05-06}
\makeatother

\begin{document}
\maketitle
\thispagestyle{empty}

\section*{Purpose}

This letter is the entry-point document for the supervisor's review of the
analytics workstream. It records the five iterations that have been closed
or parked to date, states the proposition motivating the iteration design,
and asks five methodological questions in priority order whose answers we
expect to bind the next iteration's pre-registration. A 37-page technical
companion at \texttt{docs/supervisor-review/} \texttt{2026-05-06-empirical-record-letter.pdf}
provides the per-iteration estimation detail, the verdict-mapping decision
tree, and a sample artifact for any question the supervisor wishes to
drill into. The implementation venue is on-chain (Panoptic on Uniswap v4)
but the empirical work being reviewed here is off-chain econometrics; the
on-chain primer and the full methodology are deferred to the companion.

\section{The empirical question and the iteration roster}

The project's economic motivation is best stated as a financial-intermediation
gap for an identifiable underserved population rather than as an abstract
claim about aggregate inequality. The population the iteration roster
currently centres on is Latin-American digital workers paid in local
currency (Colombian pesos, in the dev-AI Stage-1 iteration) who carry a
recurring USD-denominated cost burden through their professional inputs ---
subscription AI services, application programming interface usage fees,
cloud compute, and adjacent software-as-a-service spend. These workers are
exposed to FX-pass-through cost shocks on their monthly USD spend that no
Colombian retail-bank product currently hedges. Conventional FX-derivative
venues are inaccessible: minimum notional sizes, counterparty credit
standing, and intermediating-broker requirements are calibrated for
corporate treasury clients, not for an individual digital worker carrying a
hundred-to-low-thousands USD monthly cost line. The intermediation gap is
the operative phenomenon; aggregate distributional inequality is, under
this reading, the cumulative consequence of many such uninsured
intermediation gaps rather than the framing primitive.

The wage-to-capital boundary remains a documented empirical phenomenon and
is preserved as the analytical object of the iteration roster.
\citet{mendieta_munoz_2017} documents premature deindustrialization in
Colombia and the corresponding reallocation of labor away from
productivity-rising sectors; \citet{rodrik_2016} establishes the
cross-country pattern; \citet{mcmillan_rodrik_2011} formalises the
structural-change-and-productivity decomposition that underlies the
pattern. Together these references establish that for a substantial subset
of developing-economy populations, the path from wage income into
productive-capital ownership has become structurally narrower over the
past three decades, and that the populations on the wrong side of that
narrowing are exactly those that an absent intermediation layer fails to
insure. The project's empirical question is which macro risks \(X\) are
quantitatively responsible for the recurring uninsured cost shocks that
keep specific target populations on the wrong side of this boundary.

The unit of work is a triple \((Y, M, X)\). \(Y\) is the welfare-relevant
outcome variable on which a target population's exposure to the
wage-to-capital transition is measured (operationalised per population:
the Section-J information-and-communication employment share for young
Colombian workers, in the dev-AI Stage-1 iteration). \(X\) is the macro
risk hypothesised to block the transition for that population, identified
empirically and entered as a lag-distributed right-hand-side variable
under a pre-pinned lag structure. \(M\) is the settlement geometry of the
candidate hedging instrument on the on-chain substrate; it is declared
out of scope for empirical validation and deferred to Stage~2 of each
iteration. The validation strategy estimates a composite slope
\(\hat{\beta}_{\mathrm{composite}}\) of \(Y\) on lagged \(X\), with sign
expectation, lag structure, primary specification, and verdict thresholds
pre-registered before the panel is touched. The methodology is a
pre-registered composite-\(\beta\) OLS+HAC validation; full detail in the
companion.

As of the date of this letter, the project has run five iterations and
\emph{all five are closed or parked}: no Stage-1 work is in progress.
This letter therefore presents a complete log rather than a snapshot.

\section{Five iterations: a compact record}

\begin{table}[!htbp]
\centering
\small
\begin{tabular}{@{}p{2.6cm}p{3.4cm}p{1.7cm}p{3.0cm}p{3.4cm}@{}}
\toprule
\textbf{Population} & \textbf{(Y, X) pair} & \textbf{Verdict} &
\textbf{Headline number} & \textbf{One-line caveat} \\
\midrule
Colombian young workers (14--28), broad services & Sections G--T youth
share (logit) on \(\log(\mathrm{COP/USD})_{t-k}\), \(k \in \{6,9,12\}\)
& PASS (2026-04-28) & \(\hat{\beta}_{\mathrm{composite}} = +0.13670985\);
HAC SE \(0.0247\); \(p \approx 1.46 \times 10^{-8}\); \(N = 134\) &
Hedges correlation, not causal channel; effect concentrated at
six-month lag; sub-aggregate may be Section M, not Section J (see below).
\\ \addlinespace
Colombian young workers (14--28), Section J ICT (dev-AI Stage-1) &
Section-J youth share (logit) on \(\log(\mathrm{COP/USD})_{t-k}\),
\(k \in \{6,9,12\}\) & FAIL, sign-flipped (2026-05-06) &
\(\hat{\beta}_{\mathrm{composite}} = -0.146\); HAC SE \(0.0847\);
\(t = -1.726\); \(N = 134\) &
High-information FAIL: pre-registered \(H_1: \beta > 0\) rejected at
the negative sign. Pre-specified Section-M sensitivity arm returned
\(\hat{\beta} = +0.455\), \(p = 1.13 \times 10^{-6}\) (not
pre-registered as primary; candidate next iteration). \\ \addlinespace
Colombian corporates and remittance recipients & Weekly TRM realised
volatility on signed AR(1) CPI surprise & FAIL on weekly primary
(2026-04-19) &
\(\hat{\beta}_{\mathrm{CPI}} = -0.000685\); HAC(4) 90\,\% interval
\([-0.0036, +0.0023]\); \(N = 947\) weekly &
A1 monthly cadence and A4 release-day-excluded sensitivities both
pre-registered, both with \(90\%\) intervals strictly above zero;
A4 lower bound \(0.0005\) is small in absolute magnitude.
T1 exogeneity rejected (\(F = 15.12\)): predictive, not structural. \\
\addlinespace
Colombian household remittance recipients & Weekly TRM realised vol on
candidate on-chain remittance-flow proxy (cCOP/COPm) & EXIT\_NON\_REMITTANCE
(2026-04-24) &
No point estimates; kill-criterion exit at panel-construction stage &
0/30 peak-day flows fingerprinted as remittance; \(\sim 87\%\) of surveyed
flow decomposed into TRM arbitrage, treasury roundtripping, campaigns,
and bot activity. Kill criterion k1 fired (intent non-remittance dominant).
\\ \addlinespace
LATAM AI-policy-shock-exposed (downstream) & Bittensor SN18 (Cortex.t)
event-study apparatus on policy-event dates & PARKED (2026-04-27) &
None: apparatus complete, data ingest deferred &
Slow-lane apparatus parked to preserve intellectual capital after pivot
to fast-lane simple-\(\beta\) track (Pair~D, dev-AI Stage-1).
Re-activation conditions documented in parking memory. \\
\bottomrule
\end{tabular}
\caption{The five iterations. All numerical claims trace to \S6 of the
companion document. Pair~D, dev-AI Stage-1, and FX-vol-CPI-surprise are
econometric estimations; Phase-A.0 remittance and P1 Bittensor SN18 are
non-estimation closures. \(N_{\min} = 75\), \(\mathrm{POWER}_{\min} =
0.80\), \(\mathrm{MDES}_{\mathrm{SD}} = 0.40\) are within-iteration
pre-registration anti-fishing invariants applied uniformly across
the roster.}
\end{table}

The dev-AI Stage-1 Section-M sensitivity arm result deserves a sentence
of cross-iteration framing rather than only a per-row caveat. Read
jointly with the Pair~D PASS, the pattern --- Section J narrow-ICT
sign-flipped negative, Section M
(professional/scientific/technical/administrative services) strongly
positive on the same FX-lag panel and same monthly cadence --- suggests
that the Pair~D PASS on the broad Section-G--T cut was carried by the
Section-M-style sub-aggregate rather than by Section-J ICT-narrow tech
offshoring. The Pair~D headline numbers stand verbatim under the
original spec; what has sharpened is the interpretation of the operative
sub-aggregate. The compositional reframe is recorded in the companion
(\S6.2) and is one of the inputs to the cross-iteration multiple-comparison
question in \S\ref{sec:l5}.

The cross-iteration multiple-comparison structure is itself an open
methodological concern. Each iteration was pre-registered individually;
the project-level decisions that produced the roster (which \((Y, X)\)
pairs to spec-pin, which to drop after EXIT, which to deepen after PASS)
are not bound by an external pre-registration rule. The numerical mass at
the two tails across the five tests is asymmetric in a way that matters
for any FWER-style correction: Pair~D's one-sided
\(p \approx 1.46 \times 10^{-8}\) is overwhelmingly significant against
any reasonable \citet{benjaminihochberg1995}- or
\citet{romanowolf2005}-style correction across five iteration-level
tests; the FX-vol-CPI sensitivity arms (A4 lower bound \(0.0005\)) are
not. The full treatment is the subject of \S\ref{sec:l5}.

\section{The hedge-target proposition}

\begin{lemma}[Inequality-Differential Hedge Target]\label{lem:abrigo}
Let \(R_a(t)\) denote the return process of an asset basket representative
of capital-holding households' wealth, and let \(R_c(t)\) denote the
return process of a consumption basket representative of working-class
households' purchasing power. Define the inequality-differential process
\[
  Y_{\mathrm{inequality}}(t) \;=\; R_a(t) - R_c(t).
\]
A hedging instrument settled on \(Y_{\mathrm{inequality}}(t)\) is
welfare-relevant for working-class wealth in a way that no instrument
settled on \(R_a(t)\) alone or \(R_c(t)\) alone is, because the
differential directly indexes the relative-wealth gap that determines the
working-class household's purchasing-power decline relative to
capital-holding households.
\end{lemma}

Lemma~\ref{lem:abrigo} is stated as a proposition motivating the
iteration design, not as a proven theorem; the calibration of the
joint-leg structure remains Stage-2 work and is the subject of
\S\ref{sec:l3} below. The lemma is read here under the
parametric-cost-insurance underwriting standard
\citep{carter_etal_2017,mahul_stutley_2010,clarke_2016} rather than under
generic option-pricing standards: the working-class household holds the
policy and the differential serves as the welfare-relevant index variable
on which payouts settle.

\section{Five methodological questions}\label{sec:questions}

\subsection{Identification strategy}\label{sec:l1}

The Stage-1 specification is OLS on lagged log-FX with HAC inference.
\(\hat{\beta}_{\mathrm{composite}}\) is a predictive-regression coefficient
and we do not treat it as an impulse-response. The Pair~D closure
explicitly hedges the correlation rather than asserting a BPO causal
channel; the FX-vol-CPI iteration's T1 exogeneity test rejected at
\(F = 15.12\), confirming that exercise is also predictive rather than
structural. \emph{Question.} Should subsequent iterations move from a
predictive-regression posture to a causally-identified one --- and if so,
what identification standard does the supervisor require before declaring
a Stage-1 PASS sufficient for Stage-2 dispatch? Candidate strategies under
consideration are (i) instrumental variables, with central-bank policy-rate
decisions, oil-price shocks, or US-side AI-API price changes as candidate
instruments \citep[per][Chs.~6, 12]{hansen2022}; (ii) event-study
formulations leveraging the apparatus already developed for Phase-A.0
remittance and parked P1 Bittensor SN18; (iii) structural VAR per
\citet{kilian_lutkepohl_2017}, for which we do not currently have a
defensible identification scheme. \emph{Action we will take.} We will
adopt the supervisor's preferred standard as the binding identification
specification for the next iteration's pre-registration.

\subsection{Convex-payoff sufficiency under the
insurance-underwriting standard}\label{sec:l2}

Stage-1 estimates a mean slope. Mean-\(\hat{\beta}\) identification is
necessary but not sufficient for pricing payoffs that transmit through the
upper or lower tail of the conditional distribution rather than through
the conditional mean: this is the basis-risk question of
\citet{clarke_2016}. \emph{Question.} What empirical evidence does the
supervisor require before this \(\hat{\beta}\)-validation is sufficient to
underwrite a parametric cost-insurance product on the validated \(X\)?
Specific sub-questions: (i) is variance-amplification evidence at
\(X\)-shock dates sufficient (e.g.\ a GARCH-X specification with the
validated \(X\) entering the conditional-variance equation); (ii) is full
quantile-regression coverage of \(Y \mid X_{t-k}\) required so that the
basis-risk envelope of \citet{clarke_2016} is directly observable; (iii)
is extreme-value-theory peaks-over-threshold tail-index estimation
\citep{embrechts_kluppelberg_mikosch_1997} required to bound the
catastrophic-loss probability that
\citet{carter_etal_2017} and \citet{mahul_stutley_2010} treat as a
first-order underwriting input; (iv) should the dynamic
conditional-correlation MGARCH of \citet{engle_2002} be the standard
joint-distribution model for the differential's two legs. \emph{Action we
will take.} We will adopt the supervisor's chosen subset as the Stage-1
exit criterion for Stage-2 dispatch.

\subsection{Multi-Y differential calibration}\label{sec:l3}

Lemma~\ref{lem:abrigo}'s differential \(Y_{\mathrm{inequality}} = R_a - R_c\)
requires constructing each leg as a panel of observable component
variables. Closed iterations estimate single-\(Y\), single-\(X\)
responses; the joint-leg structure has not been calibrated and the
within-iteration thresholds were derived under a single-\(Y\) assumption.
\emph{Question.} How should the joint-leg structure be calibrated?
Specifically: (i) joint-leg correlation --- constant-correlation GARCH,
the dynamic conditional-correlation specification of \citet{engle_2002},
or a regime-switching extension; (ii) within-leg weighting ---
fixed welfare-anchored weights, principal-components weights, or
panel-estimated weights; (iii) regime-switching ---
Markov-switching mean-and-variance to accommodate the documented regime mix
(post-2014 oil collapse, COVID, Fed-tightening) noted in the Pair~D
closure, or regime indicators as exogenous covariates only; (iv) familywise
error under the multi-\(Y\) panel --- Bonferroni-Holm versus
false-discovery-rate, given that within-leg components are expected to be
highly pairwise correlated and Bonferroni is therefore likely conservative.
\emph{Action we will take.} We will adopt the supervisor's chosen joint
specification as the prerequisite for promoting Lemma~\ref{lem:abrigo}
from a motivating proposition to a calibrated estimation target.

\subsection{Anti-fishing protocol defensibility}\label{sec:l4}

The within-iteration anti-fishing invariants are
\(N_{\min} = 75\), \(\mathrm{POWER}_{\min} = 0.80\), and
\(\mathrm{MDES}_{\mathrm{SD}} = 0.40\) standard-deviation units of \(Y\).
These three thresholds were inherited from the Phase-A.0 Rev-5.3.1
specification and have not been re-derived under each subsequent
iteration's actual \(N\) and observed first-order autocorrelation
structure of the residuals. \emph{Question.} Are these three thresholds
defensible from the supervisor's standpoint, and what additional
pre-registration discipline does the supervisor require? Specifically:
(i) should the project register pre-analysis plans in a public registry
in the spirit of \citet{olken_2015} on pre-analysis plans in development
economics and \citet{christensen_miguel_2018} on transparency and
reproducibility in economics; (ii) should the project adopt the AEA
RCT-registry analogue for observational pre-registration (current
discipline pins spec-text by sha256 hash inside the project repository,
auditable internally but not externally registered); (iii) should the
three thresholds be re-derived per iteration, taking each iteration's
observed \(N\) and HAC-corrected effective sample size as inputs to the
power calculation \citep[in the spirit of][\S6.18]{hansen2022}.
\emph{Action we will take.} We will adopt the supervisor's preferred
pre-registration infrastructure for subsequent iterations, including
whether the inherited threshold triple should remain in force or be
replaced.

\subsection{Cross-iteration FWER and iteration-selection
bias}\label{sec:l5}

The five-iteration roster is closed; the cross-iteration structure is not
pre-registered. The mass at the two tails of the iteration-result
distribution is asymmetric: Pair~D's one-sided
\(p \approx 1.46 \times 10^{-8}\) survives any reasonable Bonferroni
correction across five iteration-level tests; the FX-vol-CPI-surprise
sensitivity arms (A1 monthly: \(90\%\) interval \([+0.0057, +0.0246]\);
A4 release-day-excluded: \([+0.0005, +0.0062]\)) are differentially
robust to such a correction; the dev-AI Stage-1 primary
delivered \(\hat{\beta} = -0.146\) at \(t = -1.726\), recorded as a
\emph{sign-flipped} FAIL with the realised sign at the direction of the
alternative rather than merely insignificant. \emph{Question.} What FWER
or false-discovery-rate correction regime should be applied across the
iteration history given the directional asymmetry, and how should
sensitivity-positive findings such as the FX-vol A1 and A4 results be
weighed under such a correction? In particular, should a sign-flipped
FAIL at conventional significance be counted symmetrically with the PASS
at the positive tail, or should the directional pre-registration on each
iteration shift the mass allocation? A second sub-question is
methodologically prior: iteration selection itself --- which \((Y, X)\)
pair to specify, which to drop, which to deepen --- is a multiple-testing
structure that within-iteration pre-registration does not address.
Candidate mitigations include pre-registered iteration-budget caps
committed at project initiation, an iteration-level alpha-spending
function in the spirit of group sequential designs, and pre-registered
iteration-allocation rules that do not depend on prior verdicts. A third,
sharpened sub-question: is the Pair~D PASS one test, or one test
conditional on a sub-aggregate selection? The dev-AI Stage-1 Section-M
sensitivity arm's positive estimate now suggests that the sub-aggregate
concentrating the broad-services signal is Section M rather than Section
J; if the Pair~D PASS is read with this in mind, the multiple-comparison
ledger must additionally account for an implicit search across
sub-aggregates of the broad-services panel
\citep[the selection-on-observables frame of][is one candidate
reference]{heckman_1979}. \emph{Action we will take.} We will adopt the
supervisor's preferred FWER/FDR regime, the supervisor's preferred
iteration-selection mitigation, and the supervisor's reading of the
Pair~D-conditional-on-sub-aggregate question as binding constraints on
the next iteration's design.

\section*{Closing and pointer to the companion}

Five iterations are closed or parked; no Stage-1 work is in progress.
The next iteration's pre-registration is held open pending the
supervisor's response to the five questions above. The 37-page technical
companion at
\texttt{docs/supervisor-review/2026-05-06-empirical-record-letter.pdf}
contains the full per-iteration record (companion \S6, nine-field schema
per iteration), the on-chain settlement primer (\S4), the methodology in
full (\S5, including the verdict-mapping decision tree), and four
appendices: Glossary (A), Cited Literature (B), Pair~D Verdict Header
sample artifact (C), and Power-Calculation Table (D). The supervisor is
invited to drill into any iteration's per-iteration block in companion
\S6 for the specification chain, the sha256 pre-registration governance,
and the independent-review caveats inherited from the closure memoranda.

\bibliography{refs}

\end{document}
