\documentclass[11pt,letterpaper]{article}

% --- packages (Econometrica-flavored single-column article) ---
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{array}
% microtype removed — local TeX Live lacks scalable text fonts (lmodern / cm-super / mathptmx)
% required by font-expansion; protrusion-only still triggers font issues on this install.
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{url}
\usepackage{enumitem}
\usepackage{xcolor}

\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,urlcolor=blue}

% --- math conveniences ---
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\Cov}{\operatorname{Cov}}
\newcommand{\dln}{\Delta\ln}
\newcommand{\NotionalCop}{\text{NotionalCost}^{COP}}
\newcommand{\NotionalUsd}{\text{NotionalCost}^{USD}}

\title{AI-Cost Factor Model --- Counterfactual Risk Quantification
  (Subscription-Aware Pilot, v0.2.10)\\[0.4em]
  \large Demonstration-Grade Pilot Iteration Closure}
\author{Abrigo Analytics --- Iteration \texttt{dev\_ai\_cost\_v2}}
\date{2026-05-18}

\begin{document}
\maketitle

\begin{abstract}
We report the closure of the \texttt{dev\_ai\_cost\_v2} iteration under
spec v0.2.10 (post audit-econ Delphi). On a single-user Claude Code
JSONL panel covering 2026-01-06 to 2026-05-14 (\(N = 28\) weekday rows
after first-differencing), the FX-vol contribution to AI-cost variance
is empirically approximately zero: under the conservative
covariance-attribution rule of \citet{PolitisRomano1994}, the FX share
of \(\Var(\dln\NotionalCop)\) point-estimates to \(0.0000277\) with a
\(90\%\) stationary-bootstrap confidence interval of
\([-3.4\!\times\!10^{-6},\,+4.1\!\times\!10^{-5}]\). The pre-registered
behavioral subscription-inelasticity test (R4-S3-USD), an HAC-OLS
regression of \(|\dln\NotionalUsd|\) on lagged \(|\dln\text{USDCOP}|\)
and \(|\dln\text{Tokens}|\) with \citet{NeweyWest1987} bandwidth
\(L = \lfloor T^{1/3} \rfloor = 3\), is \emph{POWER-HALTED} under the
canonical lagged-tokens partialling recipe: measured power at MDES
\(= 0.40\) residual-SD is \(0.1745\), below the
\emph{demonstration-grade} floor of \(0.50\) (\(N_{\min}=38\),
\textsc{POWER\_MIN}~\(=0.50\) --- below project defaults
\(N=75\) / \(0.80\)). The iteration is paused pending an
\(N \geq 75\) panel and makes \emph{no} population-level transmission
claim. We recommend the framework owner pivot to (i) data accumulation,
(ii) the R6 continuous-stream sibling iteration, or (iii) a target
population where the FX channel is ex-ante stronger.
\end{abstract}

\section{Introduction}
\label{sec:intro}

The Abrigo project's operating framework studies permissionless on-chain
convex hedges aimed at the Latin American wage-earner-to-capital
transition. Each iteration is a \((Y, M, X)\) triple in which
\(Y\) is an outcome surfacing the target population's exposure,
\(M\) is a Panoptic-eligible position settling the empirical risk, and
\(X\) is the major risk currently blocking the transition. The
present iteration is a \emph{candidate-X identification step}: does
FX volatility (USDCOP) materially drive AI-cost burden risk for
LATAM developers? The R5 PRIMARY arm answers a descriptive
risk-decomposition question (FX share of cost-burden variance); the
R4-S3-USD AUXILIARY arm answers a behavioral question (does USD-side
cost vol respond to FX vol despite zero marginal cost under a
subscription regime).

\paragraph{Subscription-aware proxy.}
The pilot subject (the repo maintainer) is on a Claude Pro/Max
subscription. Their realized USD cost is approximately flat per month
and FX exposure collapses to a single billing-day point. Their
\emph{notional} cost path --- rate-card price \(\times\) token usage
per the LiteLLM SHA-pinned rate table \citep{LiteLLM2025} --- serves
as a fair proxy for what a non-subscribed pay-per-API LATAM developer
would have paid. This is a strong assumption (equivalent demand under
both billing regimes) and is documented as a Section~\ref{sec:limits}
threat.

\paragraph{v0.2.10 audit-closure context.}
The v0.2.9 verdict (PARTIAL-FAIL\_TO\_REJECT under a contemporaneous
power-recipe pin) was withdrawn following an audit-econ Delphi
(2026-05-18). Four critical / high-severity amendments enter v0.2.10:
(\#4) Z-arms relabeled as \emph{same-ratio} corroborations of R5
rather than independent measurements; (\#5) demonstration-grade vs.
verdict-grade scope split with mandatory headline disclosure; (\#6/\#7)
R4-S3-COP's REGIME-CONDITIONAL FAIL removed from the corroboration
list as uninformative; (\#8) the canonical power-calculation recipe
reverted to lagged-tokens partialling, matching the regression's own
\(k=1\) lag structure. Under the lagged recipe, the power gate fires
and routes R4-S3-USD to POWER-HALT.

\section{Data}
\label{sec:data}

The panel is constructed from local Claude Code JSONL session logs on
the pilot subject's machine. Costs are computed as rate-card
\(\times\) token-counter usage across the five preserved token
categories (input, output, cache-creation-ephemeral-5m,
cache-creation-ephemeral-1h, cache-read) using the LiteLLM rate table
pinned by both commit SHA and file SHA-256 (dual-pin enforcement,
audit-econ closure \#3). USDCOP is the Banrep daily TRM. The build
window observed is 2026-01-06 to 2026-05-14.

\paragraph{Counters and parity.}
Per audit-econ closures \#1, \#9, and \#10, the row-drop bookkeeping
is split into named counters (Codex-commingling filter at parse,
non-Anthropic-model upstream filter, holiday-Monday weekday-mask
surfacing). ccusage cost-parity on the build window is
\(0.9994\!\times\) on raw cost totals; the combined-window aggregate
ratio of \(0.5386\!\times\) reflects the audit-disclosed treatment of
historical Codex commingling outside the build window and is
documented separately in \texttt{DATA\_PROVENANCE.md}.

\paragraph{Demonstration-grade pin.}
The project-level anti-fishing invariants pin \(N_{\min}=75\),
\textsc{POWER\_MIN}~\(=0.80\), \textsc{MDES\_SD}~\(=0.40\). v0.2.0
CORRECTIONS-G + CORRECTIONS-J relaxed the first two to \(38\) and
\(0.50\) without a Pair-D--style preserved-power proof. v0.2.10
Amendment~\#5 formalizes this as a separate \emph{demonstration-grade}
surface: relaxed floors are permitted, but every consumer
(verdict memo, this write-up, downstream M-design hand-off) must
disclose the relaxation in the headline. The present iteration is
demonstration-grade with \(N = 28\) observed (sub-floor relative to
the demonstration-grade \(N_{\min}=38\)). No verdict-grade claim is
made.

\section{Methodology}
\label{sec:method}

\subsection{R5 PRIMARY: descriptive counterfactual risk quantification}

The realized COP-denominated cost-burden return is the log
first-difference of \(\NotionalCop_t = \NotionalUsd_t \cdot
\text{USDCOP}_t\). Under \citet{PolitisRomano1994} stationary
bootstrap with expected block length
\(\lceil T^{1/3} \rceil\) and \(B = 10{,}000\) replicates, we report
the variance decomposition with \emph{conservative covariance
attribution} (CORRECTIONS-Q):
\begin{equation}
\Var(\dln\NotionalCop) \;=\; \Var(\dln\NotionalUsd) \;+\;
\Var(\dln\text{USDCOP}) \;+\; 2\,\Cov(\dln\NotionalUsd,\,
\dln\text{USDCOP}).
\label{eq:decomp}
\end{equation}
The headline FX share is
\(\Var(\dln\text{USDCOP}) / \Var(\dln\NotionalCop)\) \emph{excluding}
the covariance term, which is reported separately as a diagnostic
with its own bootstrap CI. The covariance-exclusion rule is
conservative for the M-design gate: positive covariance under
alternative attribution schemes would inflate the apparent FX share.

\paragraph{R5 verdict logic.}
None. R5 is descriptive --- it produces a number with bootstrap CI
consumed by the M-design step. The only pre-pinned quality threshold
is the CI half-width on the FX share \(\leq 0.15\), an M-design
usability gate (CORRECTIONS-T), not a verdict gate.

\subsection{R4-S3-USD AUXILIARY: behavioral subscription-inelasticity}

The R4-S3-USD specification is
\begin{equation}
|\dln\NotionalUsd_t| \;=\; \alpha_0^{USD} \;+\; \alpha_1^{USD}\,
|\dln\text{USDCOP}_{t-k}| \;+\; \alpha_2^{USD}\,
|\dln\text{Tokens}_{t-k}| \;+\; u_t^{USD},
\label{eq:r4s3usd}
\end{equation}
with \(k=1\) primary and \(k=5\) sensitivity. The pre-pinned null is
\(\alpha_1^{USD}=0\) under a \emph{two-sided} test (CORRECTIONS-S):
under the subscription regime, marginal USD cost is zero, so the
behavioral channel should produce no response in either direction;
rejection either way is the informative finding. HAC standard errors
follow \citet{NeweyWest1987} with the
\citet[Andrews]{Andrews1991}-style data-dependent bandwidth
\(L = \lfloor T_{\text{lag}}^{1/3} \rfloor\), automated per
\citet{NeweyWest1994}.

\paragraph{Power-HALT checkpoint.}
Power is Monte-Carlo'd at MDES \(=0.40\) residual-SD on the
post-control residual variance of \(|\dln\NotionalUsd|\) after
partialling \(|\dln\text{Tokens}|\). Per CORRECTIONS-U the
power-HALT was expected to fire at \(T = 38\). The canonical
partialling-regressor recipe is \emph{lagged tokens} (matching the
regression's own \(k=1\) lag structure), per v0.2.10 Amendment \#8;
the contemporaneous-tokens recipe is reported for transparency but
does not gate the verdict.

\subsection{Z-arms sensitivity (diagnostic only)}

Z-1 aggregates the same variance ratio at weekly (Z-1a) and monthly
(Z-1b) horizons. Z-2-main back-casts the observed \(\dln\NotionalUsd\)
distribution onto the 2024--2025 daily TRM via stationary bootstrap
(\(B_{\text{paths}} = 1{,}000\)); Z-2-null repeats the loop with the
2026-Q1-Q2 TRM only for null calibration; Z-2-W conditionally
activates a Winsorized robustness sub-arm if observed excess kurtosis
exceeds \(3.0\). Z-3 is the escalation gate
(OR over relative-\(5\!\times\)-baseline and absolute-\(0.05\)
thresholds, with null-calibration suppression). \textbf{Per v0.2.10
Amendment \#4, all Z-arms decompose the same algebraic variance ratio
across different horizons / TRM windows and are
\emph{same-ratio corroborations} of R5, not independent
measurements.}

\section{Results}
\label{sec:results}

\subsection{R5 PRIMARY}

The headline FX share over the daily 2026-Q1-Q2 window
(\(T = 29\) daily rows, \(N = 28\) post-first-diff) is reported in
Table~\ref{tab:r5}. The FX share's stationary-bootstrap CI half-width
is \(2.2 \!\times\! 10^{-5}\), three orders of magnitude inside the
\(0.15\) M-design usability threshold. The covariance term is small,
negative (\(-0.002924\)), and its CI crosses zero --- consistent with
a weak natural-hedge effect within the developer's behavior, but not
significant.

\begin{table}[h]
\centering
\caption{R5 PRIMARY variance decomposition --- conservative attribution
  (CORRECTIONS-Q). 90\% stationary-bootstrap CI,
  \citet{PolitisRomano1994}, \(B = 10{,}000\), expected block length
  \(\lceil T^{1/3} \rceil = 4\).}
\label{tab:r5}
\begin{tabular}{lccc}
\toprule
Component & Point estimate & 90\% CI lower & 90\% CI upper \\
\midrule
FX share (excl.\ cov)              & \(0.0000277\)  & \(-3.4 \times 10^{-6}\) & \(+4.1 \times 10^{-5}\) \\
Usage share (excl.\ cov)           & \(1.001078\)   & \(0.998897\)            & \(1.003730\) \\
\(2 \times \Cov\) (diagnostic)     & \(-0.002924\)  & \(-0.008650\)           & \(+0.002815\) \\
\midrule
\(\hat\sigma_{\dln\NotionalCop}\)  & \(1.659\)      & \(1.216\)               & \(2.125\) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{R4-S3-USD: POWER-HALT}

The R4-S3-USD primary regression at \(k=1\) yields
\(\hat\alpha_1^{USD} = -18.116\), HAC SE \(= 40.221\),
\(t = -0.450\), two-sided \(p = 0.6524\), with HAC bandwidth
\(L = \lfloor 28^{1/3} \rfloor = 3\). \textbf{These coefficients are
reported for the record only; they do not adjudicate the null.}
Under the canonical lagged-tokens partialling recipe, the measured
power at MDES \(= 0.40\) residual-SD is \(0.1745\), below the
demonstration-grade floor of \(0.50\). The contemporaneous-tokens
recipe (diagnostic only) yields \(0.7115\). The verdict label is
\textbf{POWER-HALT}; PARTIAL-REJECT and PARTIAL-FAIL\_TO\_REJECT are
\emph{not applicable} because the power gate fires before the
\(N\)-floor gate. Table~\ref{tab:r4s3usd} summarizes.

\begin{table}[h]
\centering
\caption{R4-S3-USD HAC-OLS results (\(k=1\) primary, \(k=5\)
  sensitivity) and power gate. Reported coefficients are for-the-record
  only; the verdict is POWER-HALT.}
\label{tab:r4s3usd}
\begin{tabular}{lcc}
\toprule
                                         & \(k=1\) (primary)   & \(k=5\) (sensitivity) \\
\midrule
\(\hat\alpha_1^{USD}\)                   & \(-18.116\)         & \(-35.750\) \\
HAC SE                                   & \(40.221\)          & --- \\
\(t\)-statistic                          & \(-0.450\)          & --- \\
Two-sided \(p\)                          & \(0.6524\)          & --- \\
HAC bandwidth \(L\)                      & \(3\)               & \(2\) \\
\midrule
Residual SD (lagged partialling)         & \multicolumn{2}{c}{\(1.156\)} \\
MDES effect \(0.40 \times \sigma_e\)     & \multicolumn{2}{c}{\(0.462\)} \\
Power (lagged, \textbf{canonical})       & \multicolumn{2}{c}{\(\mathbf{0.1745}\) (gate: \textbf{FAIL})} \\
Power (contemporaneous, diagnostic)      & \multicolumn{2}{c}{\(0.7115\) (gate: pass)} \\
\midrule
\textbf{Verdict}                         & \multicolumn{2}{c}{\textbf{POWER-HALT}} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Z-arms (diagnostic, same-ratio)}

Table~\ref{tab:zarms} reports the same-ratio Z-arms. The Z-3 verdict
is \texttt{NOT\_ESCALATE}: no Z-arm trips the OR-rule
(\(\geq 5\!\times\) baseline or \(\geq 0.05\) absolute). Per v0.2.10
Amendment \#4, these arms are \emph{not} independent measurements
and are not counted as corroborating evidence for the headline R5
finding.

\begin{table}[h]
\centering
\caption{Z-arms (same-ratio corroborations of R5; diagnostic-only).}
\label{tab:zarms}
\begin{tabular}{lcc}
\toprule
Arm                                  & FX share        & Notes \\
\midrule
Daily baseline (R5)                  & \(0.0000277\)   & primary \\
Z-1a weekly                          & \(0.000073\)    & \(T_w \approx 9\) \\
Z-1b monthly                         & \(0.000099\)    & \(T_m = 4\), TAGGED uninformative \\
Z-2-main backcast median             & near baseline   & 2024--2025 TRM, \(B_{\text{paths}}=1{,}000\) \\
Z-2-W (Winsorized)                   & conditional     & activates iff excess kurtosis \(> 3.0\) \\
\midrule
Z-3 verdict                          & \multicolumn{2}{c}{\texttt{NOT\_ESCALATE}} \\
\bottomrule
\end{tabular}
\end{table}

\section{Discussion}
\label{sec:discuss}

\paragraph{Regime-conditional reading.}
For this user \(\times\) this 2026-Q1-Q2 window, individual
cost-burden volatility is dominated by token-burst usage variance,
not by FX moves. The usage-share point estimate of \(1.001\) (with
its CI tightly around \(1.0\)) and the FX-share point of
\(2.77 \!\times\! 10^{-5}\) (CI containing zero on the lower end)
together imply that the cost panel's variance, on this window, is
algebraically explained by USD-side variance with a small negative
covariance adjustment. The reading is descriptive, regime-conditional,
and bounded to the n=1 proxy subject.

\paragraph{What R5 does and does not say.}
R5 is Role A (individual-sizing) per CORRECTIONS-R. The implication
for the M-design step is that an individual hedge for this user, at
this notional burn rate, should be dominated by token-volume risk
exposure rather than FX exposure. \emph{This is not a population-level
channel-existence claim.} Channel existence under the framework
requires the R4-S3-USD behavioral test, which here POWER-HALTs.

\paragraph{Why R4-S3-USD cannot be adjudicated.}
The behavioral channel asks whether a subscription user nevertheless
shifts token consumption in response to FX volatility (e.g.,
defensively reducing usage when COP weakens, despite the flat USD
invoice). At \(N=28\) and HAC-corrected SE of \(40.2\) on
\(\hat\alpha_1^{USD}\), the regression cannot distinguish meaningful
behavioral response from noise. The high \(p = 0.6524\) is
\emph{uninformative} once the power gate has failed: with \(B=2{,}000\)
simulated draws at MDES \(0.40 \sigma_e\), only \(17\%\) of draws
reject under the lagged canonical recipe. The honest reading is
"insufficient data," not "no effect."

\paragraph{Why Z-arms cannot rescue R4-S3-USD.}
Z-arms re-decompose the same R5 ratio across alternate time horizons
and TRM windows. They corroborate the descriptive small-FX-share
finding (regime-conditional) but carry no independent inferential
weight on the behavioral channel. Counting them as independent
evidence is fishing-by-redundancy under \S 7's anti-fishing
invariants.

\section{Limitations}
\label{sec:limits}

\begin{enumerate}[leftmargin=*,nosep]
\item \textbf{Sub-floor \(N\).} Observed \(N = 28\) is below the
  demonstration-grade \(N_{\min} = 38\) (itself below the project
  default of \(75\)). All findings are pilot-grade.
\item \textbf{Single-user pilot.} The panel is a single-user JSONL
  trace. No population-level inference is admissible.
\item \textbf{Subscription-as-proxy.} The notional-cost path
  substitutes for what a non-subscribed pay-per-API LATAM developer
  would have paid for equivalent usage. Equal demand under both
  billing regimes is a strong assumption.
\item \textbf{v0.2.10 audit-econ corrections context.} The withdrawn
  v0.2.9 verdict (PARTIAL-FAIL\_TO\_REJECT under the contemporaneous
  power-recipe pin) remains in git history for audit trail. The
  current write-up reports the lagged canonical recipe per
  Amendment \#8.
\item \textbf{Regime-conditional reading.} The small-FX-share
  finding is conditional on the 2026-Q1-Q2 USDCOP regime and on the
  pilot subject's usage pattern. Generalization is not claimed.
\item \textbf{Same-ratio corroboration discipline.} Z-arms are not
  independent inferential evidence; they are reported as
  diagnostic-only.
\end{enumerate}

\section{Conclusion}
\label{sec:conclusion}

The \texttt{dev\_ai\_cost\_v2} iteration is paused at
\textbf{demonstration-grade pending an \(N \geq 75\) panel} ---
explicitly not closed-PASS and explicitly not closed-FAIL. The R5
descriptive finding (FX share \(\approx 0\%\), regime-conditional)
informs M-design priors under Role A but is not a framework-level
channel-existence statement. The R4-S3-USD behavioral test
POWER-HALTs; it cannot be adjudicated at current sample size and
power.

Three pre-enumerated pivots are available to the framework owner:
(i) continue accumulating data on the pilot subject (\(\sim 12\)--\(18\)
months to reach \(N \geq 75\) at observed cadence; regime-change risk
material); (ii) pivot within the same \(Y\) to the R6 sibling
iteration on continuous-stream simulation, parked at v0.1.3 and ready
for dispatch; or (iii) pivot to a different \(Y\) (target population)
where the FX channel is ex-ante stronger. The choice is the framework
owner's. None of the three implies that the FX-channel hypothesis has
been disconfirmed; only that this particular pilot, on this
subscription user and this window, cannot adjudicate it.

\begin{thebibliography}{9}

\bibitem[Andrews(1991)]{Andrews1991}
Andrews, D.\,W.\,K. (1991). Heteroskedasticity and Autocorrelation
Consistent Covariance Matrix Estimation. \emph{Econometrica} 59(3),
817--858.

\bibitem[LiteLLM(2025)]{LiteLLM2025}
LiteLLM. (2025). Model price and context-window registry.
\url{https://github.com/BerriAI/litellm} (commit SHA + file SHA-256
pinned in DATA\_PROVENANCE.md).

\bibitem[Newey and West(1987)]{NeweyWest1987}
Newey, W.\,K., and K.\,D. West (1987). A Simple, Positive
Semi-Definite, Heteroskedasticity and Autocorrelation Consistent
Covariance Matrix. \emph{Econometrica} 55(3), 703--708.

\bibitem[Newey and West(1994)]{NeweyWest1994}
Newey, W.\,K., and K.\,D. West (1994). Automatic Lag Selection in
Covariance Matrix Estimation. \emph{Review of Economic Studies}
61(4), 631--653.

\bibitem[Politis and Romano(1994)]{PolitisRomano1994}
Politis, D.\,N., and J.\,P. Romano (1994). The Stationary Bootstrap.
\emph{Journal of the American Statistical Association} 89(428),
1303--1313.

\end{thebibliography}

\end{document}
