\section{Introduction \& Economic Motivation}\label{sec:1}

This memorandum is an empirical-methodology record and a request for
guidelines. It documents the iteration history that the analytics workstream
has executed to date, presents the methodology that produced each verdict,
and asks five prioritised methodological questions whose answers we expect to
bind the design of subsequent iterations. The memorandum is addressed to the
analytics supervisor; we expect the supervisor's response, where given, to
serve as the design constraint for the next iteration's pre-registration.

The project's economic motivation is best stated as a financial-intermediation
gap for an identifiable underserved population rather than as an abstract
claim about aggregate inequality. The population the iteration roster
currently centres on is Latin-American digital workers paid in local
currency (Colombian pesos, in the dev-AI Stage-1 iteration of
Section~\ref{sec:7}) who carry a recurring USD-denominated cost burden
through their professional inputs --- subscription AI services, application
programming interface usage fees, cloud compute, and adjacent
software-as-a-service spend. These workers are exposed to FX-pass-through
cost shocks on their monthly USD spend that no Colombian retail-bank
product currently hedges. Conventional FX-derivative venues are
inaccessible: minimum notional sizes, counterparty credit standing, and
intermediating-broker requirements are calibrated for corporate treasury
clients, not for an individual digital worker carrying a
hundred-to-low-thousands USD monthly cost line. The intermediation gap is
the operative phenomenon; aggregate distributional inequality is, under
this reading, the cumulative consequence of many such uninsured
intermediation gaps rather than the framing primitive.

We acknowledge that the project's intellectual origin is post-Keynesian and
that the iteration roster is descended from work that framed the problem in
distributional terms. The present document does not litigate that
intellectual provenance against the supervisor's neoclassical training.
Instead we treat the framework's contribution as the identification and
empirical validation of the macro risks \(X\) that produce the cost shocks,
followed by the design of \emph{parametric cost-insurance products} that
close the intermediation gap. This framing is deliberate: it places the
methodology under the standards of the parametric-index-insurance
literature \citep{carter_etal_2017,mahul_stutley_2010,clarke_2016} rather
than under the standards of generic derivatives pricing, and it lets the
supervisor evaluate the empirical work using the underwriting discipline he
would apply to any insurance product (basis risk, adverse selection,
catastrophic-loss probability, counterparty matching). The implementation
venue described in Section~\ref{sec:4} is on-chain settlement
infrastructure; that infrastructure is not the economic primitive.

The wage-to-capital boundary remains a documented empirical phenomenon and
is preserved as the analytical object of the iteration roster.
\citet{mendieta_munoz_2017} documents premature deindustrialization in
Colombia and the corresponding reallocation of labor away from
productivity-rising sectors; \citet{rodrik_2016} establishes the
cross-country pattern; \citet{mcmillan_rodrik_2011} formalises the
structural-change-and-productivity decomposition that underlies the
pattern. Together these references establish that for a substantial subset
of developing-economy populations, the path from wage income into
productive-capital ownership has become structurally narrower over the past
three decades, and that the populations on the wrong side of that
narrowing are exactly those that an absent intermediation layer fails to
insure. The project's empirical question is which macro risks \(X\) are
quantitatively responsible for the recurring uninsured cost shocks that
keep specific target populations on the wrong side of this boundary.

The unit of work is a triple \((Y, M, X)\) defined formally in
Section~\ref{sec:2} and motivated by a hedge-target proposition in
Section~\ref{sec:3}. The validation strategy estimates a composite slope
\(\hat{\beta}_{\text{composite}}\) of \(Y\) on lagged \(X\), with sign
expectation, lag structure, primary specification, and verdict thresholds
pre-registered before the panel is touched. Section~\ref{sec:5} (drafted by
the data-methodology agent) presents the methodology in full; we mention it
here only to fix the object the supervisor is being asked to review.

\paragraph{Iteration count and the cross-iteration multiple-comparison
posture.} The project has run five iterations to date: one PASS verdict
(Pair~D, BPO offshoring against COP/USD lag, closed 2026-04-28); one
in-progress (dev-AI Stage-1, Section~J ICT employment share against
COP/USD lag, dispatched 2026-05-05); one closed FAIL on the pre-registered
weekly primary with two pre-registered sensitivities producing positive
confidence intervals (FX-vol-on-CPI-surprise); one closed
\texttt{EXIT\_NON\_REMITTANCE} (Phase-A.0 remittance corridor, closed
2026-04-24); and one parked-for-record (P1 Bittensor SN18 event-study
apparatus). Each individual iteration was pre-registered, and the
within-iteration anti-fishing discipline is documented in
Section~\ref{sec:5}. The \emph{cross-iteration} multiple-comparison
structure, however, is not addressed by within-iteration pre-registration:
which \((Y, X)\) pairs entered the iteration roster in the first place,
which were dropped after \texttt{EXIT}, and which were deepened after
\texttt{PASS} are decisions made by the project team, not by a
pre-registered iteration-allocation rule. We acknowledge that this
constitutes an open methodological question and we treat it explicitly in
Section~\ref{sec:7} (specifically Section~\ref{sec:7.5}) rather than
presenting the five verdicts as if they were drawn from independent,
pre-iteration-selected experiments. The supervisor's guidance on the
appropriate cross-iteration familywise-error or selection-bias mitigation
regime is one of the five asks in Section~\ref{sec:7}.

The remainder of the memorandum is organised as follows.
Section~\ref{sec:2} defines the operating framework and the \((Y, M, X)\)
triple. Section~\ref{sec:3} introduces the inequality-differential hedge
proposition that motivates the iteration design and reads it as the
underwriting target of a parametric cost-insurance product. The on-chain
settlement substrate is summarised by a separate agent in
Section~\ref{sec:4} for completeness; it is contextual rather than
evidentiary, and the supervisor need not engage with it to evaluate the
econometrics. Sections~\ref{sec:5} and~\ref{sec:6} present the methodology
and the iteration log. Section~\ref{sec:7} states the five methodological
questions in priority order. Section~\ref{sec:8} closes with the request
for guidelines.
