\section{The Abrigo Lemma: Inequality Differential as Hedge Target}\label{sec:3}

This section states the proposition that motivates the choice of \(Y\) in
the framework's medium-term iteration roster. It is presented as a
proposition rather than a proven theorem; the disclaimer is repeated
immediately after the formal statement and is binding on the rest of the
memorandum. The economic interpretation given to the proposition is that
of a parametric-cost-insurance underwriting target rather than that of a
generic derivative hedge target; the mathematical content is unchanged
from earlier statements internal to the project, but the framing under
which the supervisor is invited to evaluate it is the
parametric-index-insurance literature
\citep{carter_etal_2017,mahul_stutley_2010,clarke_2016}.

\begin{lemma}[Inequality-Differential Hedge Target]\label{lem:abrigo}
Let \(R_a(t)\) denote the return process of an asset basket representative
of capital-holding households' wealth, and let \(R_c(t)\) denote the
return process of a consumption basket representative of working-class
households' purchasing power. Define the inequality-differential process
\[
  Y_{\text{inequality}}(t) \;=\; R_a(t) - R_c(t).
\]
A hedging instrument settled on \(Y_{\text{inequality}}(t)\) is
welfare-relevant for working-class wealth in a way that no instrument
settled on \(R_a(t)\) alone or \(R_c(t)\) alone is, because the
differential directly indexes the relative-wealth gap that determines the
working-class household's purchasing-power decline relative to
capital-holding households.
\end{lemma}

\paragraph{Status of the Lemma --- load-bearing disclaimer.} Lemma
\ref{lem:abrigo} is stated as a proposition motivating the project's
iteration design. It is \emph{not} a proven theorem. In particular, the
existence of a settlement-feasible instrument that materialises the
differential \(Y_{\text{inequality}}\) on the on-chain substrate, and the
calibration of the associated stochastic process, are Stage-2 / Stage-3
work that this memorandum does \emph{not} prove. The proposition's role
in the memorandum is solely to motivate why the iteration roster targets
distributional differentials rather than single-aggregate macro
variables.

\paragraph{Insurance-framing reading.} Under the framing developed in
Section~\ref{sec:1}, the differential \(Y_{\text{inequality}}(t)\) is the
welfare-relevant index variable on which a parametric cost-insurance
product should settle. The working-class household holds the policy ---
the \emph{insured} party --- and the policy pays out as a function of the
realised differential. The capital-holding counterparty already carries
the asset exposure represented by \(R_a(t)\) through pre-existing
portfolio positions and therefore does not need to be matched to a new
opt-in derivative leg; the counterparty matching is structural rather
than negotiated, in the sense documented for parametric-index insurance
by \citet{carter_etal_2017} and surveyed for developing-country
agricultural applications in \citet{mahul_stutley_2010}. The basis-risk
posture that \citet{clarke_2016} formalises --- the gap between an
index-payoff and the policy-holder's realised loss --- becomes the
appropriate Stage-1 evaluation discipline, and that posture is the one
under which the supervisor is asked to evaluate the empirical work in
Section~\ref{sec:7.2}.

\subsection{Why a single-aggregate hedge is insufficient}

A standard single-aggregate macro hedge --- for example, an instrument
settled on the headline consumer-price index alone, or on the bilateral
exchange rate alone --- fails to capture the relative-wealth dimension
that matters for working-class welfare. The headline consumer-price index
mixes consumption baskets across the wealth distribution; capital-holding
households frequently hold consumer-price-index-indexed assets that
neutralise their own exposure to inflation, leaving the working-class
household systematically more exposed than the headline number suggests.
A hedge settled on the headline index therefore prices off the
\emph{average} household's exposure rather than off the \emph{relative}
exposure differential that is welfare-relevant for the population of
interest. Read in insurance terms, the headline-index policy carries
unmodelled basis risk against the policy-holder's actual loss exposure;
\citet{clarke_2016} establishes that uninternalised basis risk on an
index-payoff product is a first-order determinant of rational demand for
that product. The differential
\(Y_{\text{inequality}}(t) = R_a(t) - R_c(t)\) removes the aggregation
problem by construction: it indexes precisely the gap between the two
baskets rather than the level of either, and therefore admits a
basis-risk treatment of the kind \citet{clarke_2016} requires.

\subsection{Two-sided market and counterparty matching}

A second motivation for the differential formulation is the existence of
a natural two-sided market that does not require an opt-in negotiated
counterparty. On the differential, the working-class household is
naturally LONG (a widening differential corresponds to a relative loss
for the household, which the policy-payoff compensates), while
capital-holding households are naturally SHORT through the asset
positions they already hold (their existing portfolios deliver the
\(R_a(t)\) leg, so their natural exposure is to a narrowing
differential). The differential formulation therefore admits the same
counterparty-matching structure that \citet{carter_etal_2017} characterise
for index-insurance arrangements: the index payoff clears the
counterparty-availability problem that obstructs single-aggregate
constructions, because the counterparty is not negotiated into a new
position but is structurally exposed to the index already. The
counterparty-matching argument is conjectural at the present stage; we
flag the formal calibration of the differential's joint-leg correlation
structure and FWER posture as the methodological question in
Section~\ref{sec:7.3}.

\subsection{Calibration of the continuous-time differential}

The differential \(Y_{\text{inequality}}(t)\) is treated as a
continuous-time stochastic process. Its calibration --- choice of the
component variables that constitute each leg, weighting within each leg,
joint-leg correlation structure, and treatment of regime switches --- is
deferred to Section~\ref{sec:7.3} as an explicit ask for the supervisor's
input. The project's current iteration roster has not yet attempted full
joint calibration; closed and active iterations estimate the response of
single \(Y\) variables to single \(X\) lags as a precursor exercise. The
parametric-cost-insurance reading developed above does not change the
mathematical object being calibrated; it changes the standard against
which the calibration's adequacy is judged, and that standard is the
underwriting-discipline standard developed across
\citet{carter_etal_2017}, \citet{mahul_stutley_2010}, and
\citet{clarke_2016}.
