\section{On-Chain Settlement Substrate: Primer for the Non-Specialist Reader}\label{sec:4}

This section introduces the on-chain settlement venue at the level required for
the supervisor to evaluate the empirical record presented in Sections~\ref{sec:5}
and~\ref{sec:6}. We anchor each on-chain construct to its closest
traditional-finance analogue at first introduction and avoid implementation
detail throughout. We emphasise at the outset that the empirical
\(\hat{\beta}\)-validation work the supervisor is reviewing is conducted on
off-chain data series (DANE, Banrep, GEIH); on-chain content is contextual,
not evidentiary.

\subsection{What ``On-Chain'' Means}\label{sec:4-onchain-meaning}

In the present document, an instrument is said to settle \emph{on-chain} when
the rule-set governing its cash flows --- the conditions under which obligations
are recognised, netted, and discharged --- is recorded on a public ledger and
executed by a programmable settlement engine. The closest traditional-finance
analogue is a \emph{clearinghouse with a public, verifiable, programmable
rule-set}: the contractual specification that a clearinghouse keeps in its
internal rulebook is, in this setting, a publicly inspectable object whose
execution is automatic and observable to all market participants.

Two properties of this venue are load-bearing for what follows. First,
settlement is \emph{deterministic} given the rule-set and the observable inputs;
there is no discretionary intermediary that must be coordinated with at trade
time. Second, the rule-set is \emph{programmable in advance}: payoff geometries
that would conventionally require a bespoke ISDA confirmation can be encoded once
and reused at scale. We emphasise that nothing in this section depends on the
implementation language of the rule-set; the reader is asked only to accept that
a programmable, publicly-auditable clearinghouse exists as the settlement
backdrop for the instrument family the framework targets.

\subsection{Panoptic: Perpetual Options on Liquidity-Provider Positions}\label{sec:4-panoptic}

Panoptic is a deployed protocol that issues perpetual option-like instruments
written against concentrated liquidity-provider (LP) positions on the Uniswap
\textsc{v3}/\textsc{v4} automated market makers
\citep{panoptic_whitepaper_2023,panoptic_docs_2026,uniswap_v3_whitepaper_2021,uniswap_v4_whitepaper_2024}.
The traditional-finance analogue is a \emph{perpetual-future-on-options
instrument} in which the option strike is encoded as the price range over
which a market-making position is concentrated, and the option premium accrues
continuously as a streaming funding payment rather than as a one-shot premium
at trade inception.

The economic content can be stated without recourse to implementation. A
concentrated LP position, in the Uniswap \textsc{v3}/\textsc{v4} sense, is a
two-asset market-making commitment over a bounded price interval
\([P_{\ell}, P_{u}]\). Inside the interval, the position behaves as a passive
market maker; outside, it converts entirely to one of the two assets. The
payoff of such a position, viewed as a function of the terminal price \(P_T\),
is a piecewise-concave curve --- a long-gamma exposure inside the interval and
a directional position outside it. Panoptic's contribution is to allow market
participants to take \emph{either side} of this payoff continuously: a short
position in the LP curve (sold to a long counterparty for a streaming premium)
delivers a payoff geometry equivalent to a long perpetual option, with the
strike, range, and convexity controlled by the chosen \([P_{\ell}, P_{u}]\)
bracket and the asset pair. Standard option structures --- covered calls,
range LPs, perpetual puts, and straddles --- are all expressible as
combinations of such positions.

For the framework's purposes, three properties matter. First, the instruments
are \emph{convex} in the underlying price: convexity is the load-bearing
property required for the (Y, M, X) instrument family articulated in
Section~\ref{sec:2} and the inequality-differential lemma of
Section~\ref{sec:3}. Second, the instruments are \emph{perpetual}: they do
not expire on a fixed calendar date and therefore admit indefinite
roll-without-rebooking, which is operationally relevant for a wage earner
paying a small recurring premium out of wage income. Third, the instruments
are \emph{permissionless}: any market participant who can interact with the
public ledger can take either side, without bilateral counterparty
negotiation. We do not litigate liquidity adequacy at the venue here;
Section~\ref{sec:4-disclaimer} returns to the question of how the framework
treats the gap between empirical \(\hat{\beta}\)-validation and live deployment
liquidity.

\subsection{Mento Purchasing-Parity Stablecoins}\label{sec:4-mento}

Mento is a deployed reserve-backed stablecoin protocol on the Celo public
ledger \citep{mento_protocol_overview_2026,mento_v3_deployments_2026}. As of
the manifest dated 2026-04-27, the Mento V3 deployment maintains a roster of
twelve country-currency-pegged claims, of which the framework treats five as
load-bearing for the target populations under iteration: \textsc{usdm} (United
States dollar), \textsc{eurm} (Euro), \textsc{brlm} (Brazilian real),
\textsc{kesm} (Kenyan shilling), and \textsc{copm} (Colombian peso). Each
claim is, for the supervisor's purposes, a \emph{digital country-currency-pegged
liability of the protocol}, with the peg maintained by an on-chain reserve
mechanism. The closest traditional-finance analogue is a narrow-purpose
digital currency board: the protocol holds a portfolio of reserve assets and
exchanges fixed amounts of the pegged claim against a basket of redemption
assets at the protocol-determined reference rate, mediated by a routing
contract referred to in the manifest as the Mento Broker.

The empirical implication for the framework is decisive and is the principal
reason Mento appears in this document at all. Consider a working-class
Colombian household contemplating a hedge against a macro risk \(X\) that
threatens its purchasing power. A hedge denominated in United States dollars
settles in a currency the household does not consume; the household then bears
residual COP/USD exchange-rate risk between hedge settlement and consumption.
A hedge denominated in \textsc{copm} settles directly in a claim pegged to
the household's consumption currency, eliminating that residual exchange-rate
exposure by construction. The hedge is, in this sense,
\emph{purchasing-parity preserving} at the unit-of-account level. The same
construction applies to \textsc{brlm} for Brazilian populations,
\textsc{kesm} for Kenyan populations, and so on across the Mento basket.

We note one item of housekeeping that may bear on supervisor questions about
on-chain attribution. There exist two distinct Colombian-peso-named claims on
the Celo ledger: the Mento-protocol-native \textsc{copm} (canonical address
\texttt{0x8A567e2a\dots4041eA}) and a separate fintech-issued
\textsc{copm-Minteo} claim (address
\texttt{0xc92e8fc2\dots1d5606}) that is not Mento-protocol-native and is
therefore out of scope for any framework-level claim about Mento user
behaviour. The audit-trail discipline of the framework treats these as
non-substitutable.

\subsection{Political-Receptivity Context}\label{sec:4-receptivity}

In late January 2025, a tariff standoff between the United States executive
and the Colombian executive produced a documented forty-eight-hour spike in
new account formation at Littio, a Colombian fintech provider of USD-pegged
stablecoin accounts; the spike was on the order of one hundred percent above
prior baseline (per the project's internal reference document
\texttt{OFFCHAIN\_COP\_BEHAVIOR.md}, which the same event-window has been used
as a structural-break event-dummy across multiple iteration specifications).
This is a \emph{user-side} responsiveness signal: Colombian retail moved into
a USD-stablecoin rail during a political dollar-access shock. We include the
observation as anecdotal context for the framework's implicit hypothesis that
working-class populations facing macro-risk shocks display measurable
willingness to engage on-chain instruments when the off-chain banking rail
becomes congested or politically encumbered. We emphasise that this
observation is \emph{not} a presidential transaction; earlier draft language
asserting that the Colombian president had publicly transacted on-chain was
unsourced and has been removed. The receptivity signal is anecdotal frame
only; it is not load-bearing for any empirical or methodological claim
elsewhere in this document.

\subsection{Critical Disclaimer: Empirical Validation is Off-Chain}\label{sec:4-disclaimer}

The framework's empirical \(\hat{\beta}\)-validation work --- the verdicts
recorded in Section~\ref{sec:6} and the methodology of Section~\ref{sec:5} ---
is conducted on off-chain data series. The Pair D PASS verdict, the
dev-AI Stage-1 in-progress estimation, the FX-vol-CPI-surprise FAIL, the
Phase-A.0 EXIT, and the P1 PARKED record all rest on data assembled from the
Departamento Administrativo Nacional de Estad\'istica (DANE), the Banco de la
Rep\'ublica de Colombia (Banrep), and the Gran Encuesta Integrada de Hogares
(GEIH). On-chain data series enter the framework only at later stages of
candidate \(X_d\) signal-design, and their role there is auxiliary and
re-pivoting in light of recent verification work
(see Section~\ref{sec:6}, dev-AI Stage-1 and Phase-A.0 EXIT entries).

The settlement venue --- Panoptic on a Mento-denominated pair --- is the
\emph{deployment target} once a (Y, X) pair passes Stage-1 empirical
validation. Stage-1 validation does not require any on-chain interaction;
Stage-2 instrument design (\(M\)-sketch) does not require live liquidity;
only Stage-3 deployment requires real liquidity-provider capital. The
supervisor's review concerns the econometrics --- the Stage-1 layer. Sections
\ref{sec:5}, \ref{sec:6}, \ref{sec:7}, and \ref{sec:8}, together with
Appendices~\ref{app:C} and~\ref{app:D}, are readable end-to-end without
further on-chain familiarity.

\subsection{Reader Takeaways}\label{sec:4-takeaways}

\begin{enumerate}
  \item On-chain perpetual options exist and are tradable today, via the
        Panoptic protocol written against Uniswap \textsc{v3}/\textsc{v4} LP
        positions \citep{panoptic_docs_2026}.
  \item Purchasing-parity stablecoin denominations exist, via the Mento
        protocol's roster of country-currency-pegged claims
        (\textsc{usdm}, \textsc{eurm}, \textsc{brlm}, \textsc{kesm},
        \textsc{copm}) \citep{mento_v3_deployments_2026}.
  \item The empirical work the supervisor is reviewing is off-chain
        econometrics on DANE, Banrep, and GEIH series.
  \item The remainder of this document does not require on-chain familiarity;
        the reader may proceed directly to Section~\ref{sec:5}.
\end{enumerate}
