\section{Empirical Methodology}\label{sec:5}

This section describes the estimation discipline that has governed every
iteration in the Abrigo log. The framework treats Stage~1 (empirical risk
validation) as a stand-alone testing exercise: an iteration either produces a
positive, statistically detectable composite coefficient on a pre-pinned lag
panel of the macro risk \(X\) for the pre-pinned outcome \(Y\), or it does
not. The settlement geometry \(M\) on which an iteration would deploy is out
of scope for the verdict. The methodology below therefore reads as a small,
fixed kit of econometric primitives applied identically across iterations,
with a strict pre-registration discipline that fixes every nontrivial design
choice before the data are touched.

\subsection{Pre-registration discipline}\label{sec:5.1}

For every iteration we author a design specification that pins, in textual
form and \emph{before any data ingest}, the following objects: the operational
definition of \(Y\); the construction of \(X\); the lag set; the functional
form of the dependent variable; the primary inference procedure; the
robustness arms; the verdict thresholds; and the falsification tree. The
specification is hashed (sha256) at commit time and the hash is propagated
through the implementation plan, the result notebooks, and the verdict
artifact. Any post-data revision recomputes the hash and triggers a
two-wave independent review (Reality Checker plus a purpose-matched
specialist) before further execution. Within an iteration, threshold tuning
post-hoc is treated as silent fishing and is forbidden. This discipline is
in the spirit of the pre-analysis-plan literature \citep{olken2015,
christensenmiguel2018} adapted to an observational-econometrics setting in
which a public registry analogue is not yet adopted by the project; whether
to register against an external registry is one of the open questions we
address to the supervisor in Section~\ref{sec:7}.

\subsection{Anti-fishing invariants}\label{sec:5.2}

Three numerical invariants currently bind every iteration:

\begin{itemize}
  \item Minimum sample size after window compliance and lag-window loss:
    \(N_{\min} = 75\) observations.
  \item Minimum power at the pre-registered minimum detectable effect size:
    \(\mathrm{Power}_{\min} = 0.80\).
  \item Minimum detectable effect size: \(\mathrm{MDES} = 0.40\) standard
    deviations of \(Y\).
\end{itemize}

These three numbers were inherited from the Phase-A.0 specification chain
(Rev-5.3.1) and have \emph{not} been re-derived per iteration's actual
\(N\), residual autocorrelation structure, or HAC-adjusted effective sample
size. We flag this inheritance explicitly here, present whatever
back-of-the-envelope derivation is feasible at writing time in
Appendix~\ref{app:D}, and address the defensibility of the three numbers
to the supervisor as a dedicated open question in Section~\ref{sec:7}
(\(\S 7.4\)). The document is asking the supervisor whether these
thresholds are defensible by his neoclassical-econometrics lights, not
asserting that they have been validated.

\subsection{Functional form and primary inference}\label{sec:5.3}

For each iteration we test a one-sided null on the composite coefficient
across a pre-pinned lag panel:
\begin{equation}\label{eq:hypothesis}
H_0:\ \beta_{\mathrm{composite}} \leq 0
\qquad\mathrm{vs.}\qquad
H_1:\ \beta_{\mathrm{composite}} > 0,
\end{equation}
with the sign expectation grounded in the literature anchor of the
iteration's transmission chain (never inferred from the data). For the
two monthly-cadence iterations that share the COP/USD lag panel (Pair D
and dev-AI Stage-1), the primary specification is:
\begin{equation}\label{eq:primary-monthly}
Y_t^{\mathrm{logit}}
= \alpha
+ \beta_6\,\log(\mathrm{COP\_USD}_{t-6})
+ \beta_9\,\log(\mathrm{COP\_USD}_{t-9})
+ \beta_{12}\,\log(\mathrm{COP\_USD}_{t-12})
+ \varepsilon_t,
\end{equation}
where \(Y_t^{\mathrm{logit}} = \log\!\big(Y_t/(1-Y_t)\big)\) is the logit
of an interior-bounded employment share. The composite of interest is
\(\beta_{\mathrm{composite}} \equiv \beta_6 + \beta_9 + \beta_{12}\); its
hypothesis is tested via the linear-restriction variance
\(\widehat{\mathrm{Var}}(c'\hat{\beta}) = c'\widehat{\Sigma}\,c\) with
\(c = (0,1,1,1)'\), equivalent to the Wald-restriction variance and not a
delta-method object. Lags 6, 9, and 12 of the log-FX series are highly
serially correlated at monthly frequency; individual coefficients have
wide intervals, but the composite is precise because the off-diagonal
entries of \(\widehat{\Sigma}\) are negative under positive collinearity,
which deflates \(\widehat{\mathrm{Var}}(c'\hat{\beta})\) below the
sum-of-individual-variances intuition. A reading that ``individual lags
weren't significant'' is methodologically incorrect: the test is on the
composite alone.

The logit transform is justified on the standard textbook
\citep[\S 16.4]{hansen2022} grounds for an interior-bounded share whose
empirical range lies away from \(0\) and \(1\). The fractional-response
\citep{papkewooldridge1996} and beta-regression
\citep{ferraricribarineto2004} alternatives are not pre-committed because
the iterated \(Y\) ranges (Pair D \([0.55,0.75]\); dev-AI Stage-1
\([0.014,0.031]\) realized) are well-interior to \((0,1)\); a raw-share
specification is preserved as a robustness arm rather than as the primary.

\paragraph{Primary inference convention.} The primary standard error is
the homoskedastic OLS standard error per the project's anti-overclaim
discipline: HAC bandwidth selection introduces a post-data degree of
freedom that the pre-registration regime rules out from the primary tier.
HAC inference is run as the pre-pinned R4 robustness arm (see
\(\S\ref{sec:5.4}\)) at a fixed lag-truncation chosen at spec-authoring
time. Where the verdict label is invariant under the OLS and HAC standard
errors, headline numbers report the \emph{more conservative} of the two
SE conventions; where the two would diverge across the verdict boundary,
the OLS-primary SE governs the verdict and the HAC SE is reported as the
R4 row. In the closed Pair~D iteration the HAC SE produced a marginally
smaller p-value than the OLS SE, so the headline reports the larger HAC
SE as the more-conservative anti-fishing convention without re-routing
the verdict; in the closed dev-AI Stage-1 iteration, primary OLS and R4
HAC standard errors agree on the FAIL verdict label and the headline
reports the HAC SE on the same more-conservative-of-the-two convention.
The verdict label has been invariant across the two SE conventions in
every iteration that has closed to date.

\subsection{Robustness inference (HAC)}\label{sec:5.4}

The robustness HAC standard error is the Newey-West heteroskedasticity-
and-autocorrelation-consistent estimator \citep{neweywest1987} with
\emph{fixed lag truncation \(L\) chosen at spec-authoring time per
iteration}, not a data-driven automatic rule. The bandwidth values that
have been used to date are:
\begin{itemize}
  \item Monthly cadence (Pair D, dev-AI Stage-1): \(L = 12\). The
    truncation matches the longest right-hand-side lag in the spec --- the
    twelve-month offshoring contracting cycle that motivates the lag panel.
  \item Weekly cadence (FX-vol-CPI primary): \(L = 4\). The truncation was
    set at spec-authoring time on the basis of the in-sample residual
    autocorrelation diagnostic that closed at \(L = 4\) in the
    Phase-A.0 pipeline; a robustness sensitivity at \(L = 12\) is
    pre-registered as a fat-tail diagnostic, not as the primary.
\end{itemize}
Automatic-bandwidth selection rules \citep{neweywest1994, andrews1991}
are explicitly not used: a data-driven bandwidth introduces a post-data
degree of freedom that the pre-registration discipline rules out.

\subsection{Robustness arms}\label{sec:5.5}

For monthly iterations (Pair D and dev-AI Stage-1) the spec text
pre-pins four single-row alternatives, each varying exactly one design
choice from the primary; multi-dimensional re-specification (varying
two or more dimensions simultaneously) is forbidden:
\begin{itemize}
  \item R1 -- regime-dummy alternative to the methodology-break
    treatment (e.g., 2021 GEIH \emph{empalme} dummy in place of the DANE
    splice factor).
  \item R2 -- sector-cut narrowing alternative on the dependent variable
    (Pair D: Sections J + M + N as a BPO-narrow cut; dev-AI Stage-1:
    Section M as a sibling-narrow comparator to Section J).
  \item R3 -- functional-form alternative (raw share in place of logit).
  \item R4 -- HAC inference per \(\S\ref{sec:5.4}\), substituted for
    OLS standard errors.
\end{itemize}
The closed FX-vol-CPI iteration uses a wider thirteen-row pre-registered
sensitivity universe, of which the A1 monthly-cadence and A4
release-day-excluded rows are two; the universe is pinned via a
\texttt{sensitivity\_preregistration\_hash} on the panel fingerprint
artifact. Free lag tuning is forbidden in all iterations.

A coarse stability classifier maps the four-row vector to one of three
labels: AGREE (zero R-rows produce a sign-flipped composite relative to
the primary), MIXED (one or two flips), or DISAGREE (three or four
flips). DISAGREE triggers a special verdict label
SUBSTRATE\_TOO\_NOISY, regardless of the primary's significance, and
routes the iteration to a methodology-revision branch rather than to
verdict assignment. We additionally note that the R4 HAC arm is a
serial-correlation hedge and not, in itself, a heteroskedasticity hedge:
the OLS-primary inference assumes conditional homoskedasticity, and
whether to add a White / sandwich heteroskedasticity-only diagnostic as
a fifth pre-pinned arm is one of the methodology questions we put to the
supervisor in Section~\ref{sec:7} (\(\S 7.4\)).

\subsection{Persistent-regressor posture and Stambaugh
bias}\label{sec:5.5b}

The right-hand-side regressor in the monthly-cadence iterations is
\(\log(\mathrm{COP\_USD}_{t-k})\), which is empirically near-unit-root
at the monthly frequency on the realized 2014--2026 sample window. The
project's identification posture is predictive-regression in the strict
Granger-predictive sense (see \(\S\ref{sec:5.8}\)); the closed FX-vol-CPI
iteration's pre-registered T1 exogeneity test rejected at
\(F = 15.12\), \(p \approx 1.3 \times 10^{-9}\) and consequently the
project does not interpret \(\hat{\beta}\) as an impulse-response.

The predictive-regression posture, however, is known to suffer
small-sample bias when the right-hand-side regressor is persistent and
the residual is correlated with the regressor's innovation
\citep{stambaugh1999}; the bias does not vanish in finite samples even
under correct specification, and recent econometric work
\citep{phillips2014} has extended the asymptotic theory to near-unit-root
and mildly explosive regressors. The Pair~D and dev-AI Stage-1
iterations realized \(N = 134\) monthly observations, which is
modest-but-not-tiny on the conventional Stambaugh-bias scale; the
bias-correction literature suggests the order-\(1/N\) bias is small
relative to the realized HAC standard error in both iterations, but the
project has \emph{not} formally implemented a Stambaugh-corrected
estimator at writing time. This is one of the leading methodology
open-questions we put to the supervisor in Section~\ref{sec:7}
(\(\S 7.1\), identification ask, and \(\S 7.4\), invariants ask): whether
the predictive-regression posture is sufficient as Stage-1 verdict
machinery, or whether Stambaugh-correction (or equivalent
near-unit-root-aware inference) should be elevated from a deferred
methodology improvement to a pre-pinned R-arm before the next iteration
spec-pin.

\subsection{Verdict mapping}\label{sec:5.6}

The mapping is deterministic. Given the primary tuple
\((\hat{\beta}_{\mathrm{composite}}, \hat{p}_{\mathrm{one\text{-}sided}},
R\text{-consistency})\) and the residual-moment tuple
\((\hat{s}_3, \hat{s}_4)\), every leaf returns exactly one verdict:
\begin{enumerate}
  \item If the realized sample size \(N\) falls below \(N_{\min} = 75\),
    the iteration HALTs (see \(\S\ref{sec:5.7}\)).
  \item If R-consistency is DISAGREE, the verdict is
    SUBSTRATE\_TOO\_NOISY.
  \item Otherwise:
    \begin{itemize}
      \item \(\hat{\beta}_{\mathrm{composite}} > 0\) and
        \(\hat{p} \leq 0.05\) returns PASS.
      \item \(\hat{\beta}_{\mathrm{composite}} > 0\) and
        \(\hat{p} \in (0.05, 0.20]\) returns ESCALATE Clause~A.
      \item Clause~B fires when
        \(|\hat{\beta}_{\mathrm{composite}}/\widehat{\mathrm{SE}}| < 0.5\)
        \emph{and} either
        \(|\mathrm{skew}(\hat{\varepsilon})| > 1.0\) or
        \emph{Fisher excess kurtosis}
        \(\hat{s}_4 \equiv \hat{\kappa}_4 - 3 > 3.0\) (i.e., raw fourth
        standardized moment \(\hat{\kappa}_4 > 6\); a normal residual
        has Fisher excess kurtosis \(0\) and the threshold flags
        substantially leptokurtic tails). When Clause~B fires the
        verdict is ESCALATE Clause~B; otherwise FAIL.
      \item \(\hat{\beta}_{\mathrm{composite}} \leq 0\) and
        \(\hat{p} \leq 0.05\) returns FAIL with no Clause~B evaluation:
        a wrong-signed significant estimate falsifies \(H_1\), and
        convex-payoff escalation, which is grounded in the positive
        sign expectation, does not apply.
    \end{itemize}
\end{enumerate}
The numerical thresholds \(0.5\), \(1.0\), and \(3.0\) in Clause~B are
pinned at spec-authoring time and are not adjustable post-data; the
\(3.0\) threshold is on the \emph{Fisher} (excess-over-normal)
convention, not on the Pearson raw-fourth-moment convention. The
ESCALATE branches trigger a pre-authorized convex-payoff suite (a
\(\tau = 0.90\) quantile regression on the central lag, a GARCH-X
mean-equation regression, and a peaks-over-threshold extreme-value
regression on the upper-tail residuals); each suite leg has its own
one-sided p-value at \(\alpha = 0.10\) and the disjunction over the
three legs determines ESCALATE-PASS or ESCALATE-FAIL. The disjunction
structure is defended against multiple-testing critique on the grounds
that each leg estimates a distinct distributional-moment parameter that
maps to a distinct convex-instrument design, rather than three
restatements of the same parameter.

\subsection{HALT and disposition discipline}\label{sec:5.7}

When the spec text contradicts the data --- examples already encountered
include the DANE \texttt{RAMA4D\_R4} schema-stability HALT in the Pair~D
ingest and the realized cell-size and share-magnitude gap in the dev-AI
Stage-1 ingest --- the iteration HALTs by raising a typed exception. A
disposition memo is filed enumerating at least three pivot options; the
user adjudicates; a CORRECTIONS block lands in the next specification
revision documenting the disposition, the chosen pivot, and the
rationale; and a three-way independent review is run before further
execution. Auto-pivot is forbidden, and post-hoc threshold tuning is
forbidden. The discipline is documented in the project's internal
feedback memory under the \emph{pathological halt anti-fishing
checkpoint} rule.

\subsection{Identification posture}\label{sec:5.8}

Every estimate produced by the framework to date is correlation-identified
under predictive-regression posture. Wherever an exogeneity test has been
run as a diagnostic --- the FX-vol-CPI iteration ran a T1 exogeneity test
that rejected at \(F = 15.12\) and \(p \approx 1.3 \times 10^{-9}\) --- the
result has reinforced the predictive-regression reading: \(\hat{\beta}\)
describes how past information predicts current realized volatility, not
how an exogenous surprise causes future realized volatility. The
distinction is the standard Granger-predictive versus structural-causal
distinction; the document does not use \(\hat{\beta}\) as an
impulse-response. Whether to move to instrumental-variable, event-study,
or structural VAR identification before declaring a Stage-1 PASS
sufficient for Stage-2 dispatch --- and whether the persistent-regressor
posture noted in \(\S\ref{sec:5.5b}\) requires Stambaugh-bias-corrected
inference at the primary tier --- is the leading open question we put to
the supervisor in Section~\ref{sec:7}.

