\section{Open Methodological Questions for Supervisor Review}\label{sec:7}

We present five methodological questions in priority order. The first four
were enumerated at the brainstorming stage of the memorandum; the fifth
was added on Wave-2 verification feedback to make the cross-iteration
multiple-comparison concern explicit rather than allowing the supervisor
to introduce it as an unanticipated objection. For each ask, we state the
project's current posture, the references that frame the question for an
econometrician of the supervisor's training, and the specific decision
the supervisor's response would unblock.

\subsection{Identification strategy}\label{sec:7.1}

\paragraph{Current posture.} The Stage-1 specification is OLS on lagged
log-FX with HAC inference at fixed lag truncation \(L\) (see
Section~\ref{sec:5}). The estimator \(\hat{\beta}_{\text{composite}}\) is
a predictive-regression coefficient; we do not treat it as an
impulse-response. The PASS-verdict iteration (Pair~D) flags identification
non-uniqueness explicitly: the closure memorandum hedges the correlation
between Section-G--T employment share and lagged COP/USD movement and does
\emph{not} assert a causal channel running through BPO contracting. The
FAIL-verdict iteration (FX-vol-on-CPI-surprise) records a T1 exogeneity
rejection at \(F = 15.12\), \(p \approx 1.3 \times 10^{-9}\), confirming
that the FX-vol exercise is also predictive rather than structural.

\paragraph{Question for the supervisor.} Should subsequent iterations move
from a predictive-regression posture to a causally-identified one, and if
so, what identification standard does the supervisor require before
declaring a Stage-1 PASS sufficient for Stage-2 dispatch? Specifically:

\begin{itemize}
  \item Does the supervisor require an instrumental-variable strategy?
    Candidate instruments under consideration include
    monetary-policy-rate decisions of the central bank issuing the
    consumption-basket currency, oil-price shocks (relevant for
    commodity-exporting populations), and US-side AI-API price changes
    (relevant for the dev-AI Stage-1 iteration). The relevant references
    are \citet{hansen_2022}, Chapters~6 and~12, and
    \citet{wooldridge_2010}.
  \item Does the supervisor require an event-study formulation? The
    Phase-A.0 remittance work and the parked P1 Bittensor SN18 work both
    have event-study apparatus available; we have not promoted either to
    a Stage-1 primary in any closed iteration.
  \item Does the supervisor require a structural vector autoregression?
    The reference framework would be \citet{kilian_lutkepohl_2017}; the
    project does not currently have a structural identification scheme
    that we believe defensible against that reference's standards.
\end{itemize}

\paragraph{What the supervisor's answer unblocks.} A pinned identification
standard for subsequent iterations, including dev-AI Stage-1's eventual
Stage-2 dispatch.

\subsection{Convex-payoff sufficiency}\label{sec:7.2}

\paragraph{Current posture.} Stage-1 estimates a mean slope. The internal
working note
\texttt{memory/project\_abrigo\_convex\_instruments\_inequality.md}
records that mean-\(\hat{\beta}\) identification is necessary but not
sufficient for pricing payoffs that transmit through the upper or lower
tail of the conditional distribution rather than through the conditional
mean. Under the parametric-cost-insurance reading developed in
Sections~\ref{sec:1} and~\ref{sec:3}, this insufficiency is the
basis-risk question of \citet{clarke_2016}: a mean-calibrated index
specification can leave the policy-holder materially exposed in precisely
the tail states that motivate the demand for the policy in the first
place.

\paragraph{Question for the supervisor (insurance-framing).} The
operative question is: \emph{what empirical evidence does the supervisor
require before this \(\hat{\beta}\)-validation is sufficient to underwrite a
parametric cost-insurance product on the validated \(X\)?} We adopt the
parametric-index-insurance literature
\citep{carter_etal_2017,mahul_stutley_2010,clarke_2016} as the
methodological reference rather than the generic option-pricing
literature, on the deliberate ground that the insurance literature
imposes the stricter evidentiary standard. Where option-pricing
methodology is concerned principally with the second moment of the
underlying, the parametric-insurance methodology requires the
underwriter to characterise basis risk against the policy-holder's
realised loss \citep{clarke_2016}, the adverse-selection structure of the
policy-holder pool \citep{mahul_stutley_2010}, and the catastrophic-loss
probability and reinsurance posture of the product
\citep{carter_etal_2017,mahul_stutley_2010}. Under that stricter
standard, the specific Stage-1 questions become:

\begin{itemize}
  \item Is variance-amplification evidence at the \(X\)-shock dates
    sufficient (for example, a GARCH-X specification with the validated
    \(X\) entering the conditional-variance equation), or is variance
    amplification merely necessary?
  \item Is full quantile-regression coverage required, with explicit
    estimation of upper- and lower-tail conditional quantiles of
    \(Y \mid X_{t-k}\), so that the basis-risk envelope of
    \citet{clarke_2016} is directly observable from the Stage-1 output?
  \item Is extreme-value-theory peaks-over-threshold tail-index estimation
    required, in the spirit of
    \citet{embrechts_kluppelberg_mikosch_1997}, in order to bound the
    catastrophic-loss probability that
    \citet{carter_etal_2017} and \citet{mahul_stutley_2010} treat as a
    first-order underwriting input?
  \item Should the dynamic conditional-correlation MGARCH specification of
    \citet{engle_2002} be adopted as the standard joint-distribution model
    for the differential's two legs in
    Section~\ref{sec:3}, given that the joint-distribution behaviour
    determines the differential-leg basis risk?
\end{itemize}

The supervisor is invited to apply whichever subset of these standards he
judges appropriate; the project's posture is to err on the side of the
stricter insurance-underwriting reading rather than the more permissive
option-pricing reading.

\paragraph{What the supervisor's answer unblocks.} The Stage-1 exit
criterion for Stage-2 dispatch under the parametric-cost-insurance
product objective, and in particular the basis-risk evidentiary standard
that Stage-1 output must meet before the validated \(X\) is treated as
underwritable.

\subsection{Multi-Y differential calibration}\label{sec:7.3}

\paragraph{Current posture.} Lemma~\ref{lem:abrigo}'s differential
\(Y_{\text{inequality}} = R_a - R_c\) requires constructing each leg as a
panel of observable component variables. Closed and active iterations
estimate single-\(Y\), single-\(X\) responses; the project has not yet
calibrated the joint-leg structure, and the within-iteration thresholds
\(\big(N_{\text{MIN}} = 75,\ \mathrm{POWER}_{\text{MIN}} = 0.80,\
\mathrm{MDES}_{\text{SD}} = 0.40\big)\) were derived under a single-\(Y\)
assumption.

\paragraph{Question for the supervisor.} How should the joint-leg
structure of \(Y_{\text{inequality}}\) be calibrated? Specifically:

\begin{itemize}
  \item Joint-leg correlation: what specification --- a constant-correlation
    GARCH-type model, the dynamic conditional-correlation specification of
    \citet{engle_2002}, or a regime-switching extension --- does the
    supervisor recommend for the conditional joint distribution of
    \((R_a, R_c)\)?
  \item Within-leg weighting: what method --- fixed welfare-anchored weights
    (the working-class consumption basket), principal-components weights,
    or weights estimated from the panel itself --- should determine the
    composition of each leg?
  \item Regime-switching: should the differential admit a
    Markov-switching mean-and-variance structure to accommodate the
    documented regime mix (post-2014 oil collapse, COVID,
    Fed-tightening) noted in the Pair~D closure memorandum, or should
    regime indicators enter as exogenous covariates only?
  \item Familywise error: under the multi-\(Y\) panel implied by the
    differential, should familywise-error-rate control proceed under
    Bonferroni-Holm, or under the false-discovery-rate framework, given
    that within-leg components are expected to be highly pairwise
    correlated and Bonferroni is therefore likely conservative?
\end{itemize}

\paragraph{What the supervisor's answer unblocks.} The joint
specification of the inequality-differential dynamics, which is a
prerequisite for promoting Lemma~\ref{lem:abrigo} from a motivating
proposition to a calibrated estimation target.

\subsection{Anti-fishing protocol defensibility}\label{sec:7.4}

\paragraph{Current posture.} The within-iteration anti-fishing
invariants are
\(N_{\text{MIN}} = 75\),
\(\mathrm{POWER}_{\text{MIN}} = 0.80\), and
\(\mathrm{MDES}_{\text{SD}} = 0.40\) standard-deviation units of \(Y\).
These three thresholds were inherited from the Phase-A.0 Rev-5.3.1
specification and have \emph{not} been re-derived under each subsequent
iteration's actual \(N\) and observed first-order autocorrelation
structure of the residuals. Appendix~D presents whatever
power-derivation transparency is feasible at the time of writing; we
flag the inheritance explicitly so that the supervisor's review is on the
actually-defensible numbers and not on the inherited ones.

\paragraph{Question for the supervisor.} Are these three thresholds
defensible from the supervisor's neoclassical-econometrics standpoint,
and what additional pre-registration discipline does the supervisor
require? Specifically:

\begin{itemize}
  \item Should the project register pre-analysis plans in a public
    registry, in the spirit of \citet{olken_2015} on
    pre-analysis plans in development economics and
    \citet{christensen_miguel_2018} on transparency and reproducibility
    in economics?
  \item Should the project adopt the AEA RCT-registry analogue for
    observational pre-registration? The current discipline pins
    spec-text by sha256 hash inside the project repository; this is
    auditable internally but not externally registered.
  \item Should the three thresholds be re-derived per iteration, taking
    each iteration's observed \(N\) and HAC-corrected effective sample
    size as inputs to the power calculation, in the spirit of
    \citet{hansen_2022}, \S\,6.18 on autocorrelation-corrected effective
    sample sizes?
\end{itemize}

\paragraph{What the supervisor's answer unblocks.} The pre-registration
infrastructure for subsequent iterations, including whether the inherited
threshold triple should remain in force or be replaced.

\subsection{Cross-iteration FWER and iteration-selection bias}\label{sec:7.5}

\paragraph{Current posture.} The project has run five iterations: one
PASS, one IN PROGRESS, one FAIL on the pre-registered weekly primary with
two pre-registered sensitivities producing positive 90\,\% confidence
intervals, one \texttt{EXIT}, and one PARKED. Each iteration's
within-iteration pre-registration discipline is documented; the
\emph{cross-iteration} structure is not. Iteration-selection decisions ---
which \((Y, X)\) pairs enter the roster, which are dropped after
\texttt{EXIT}, which are deepened after PASS --- are made at the project
level and are not bound by an external pre-registration rule.

The numerical asymmetry across iterations matters here. The Pair~D
closure verdict's headline \(p \approx 1.46 \times 10^{-8}\) is
overwhelmingly significant against any reasonable Bonferroni correction
across five iteration-level tests. The FX-vol-on-CPI-surprise
sensitivities are far less robust to such a correction: the A1 monthly
sensitivity reports a 90\,\% confidence interval of
\([+0.0057,\ +0.0246]\); the A4 release-day-excluded sensitivity reports
\([+0.0005,\ +0.0062]\). The A4 lower bound at \(0.0005\) is at the edge
of significance even uncorrected, and would not survive a meaningful
familywise-error-rate adjustment.

\paragraph{Question for the supervisor.} What familywise-error-rate or
false-discovery-rate correction regime should be applied across the
iteration history, and how should sensitivity-positive findings such as
the FX-vol A1 and A4 results be weighed under such a correction?

A sub-question is methodologically prior. Iteration selection itself ---
the decision of which \((Y, X)\) pair to specify, which to drop, which
to deepen --- is a multiple-testing structure that within-iteration
pre-registration does not address. What mitigations does the supervisor
recommend? Candidate mitigations under consideration include
pre-registered iteration-budget caps committed at project initiation,
an iteration-level alpha-spending function in the spirit of group
sequential designs, and pre-registered iteration-allocation rules that
do not depend on prior verdicts. We have no current basis for choosing
among these and are explicitly asking for the supervisor's preferred
form.

\paragraph{What the supervisor's answer unblocks.} The project-level
multiple-testing discipline that supplements within-iteration
pre-registration, and the appropriate weight to assign to the
sensitivity-positive findings of the FAIL-primary iteration.
