% =========================================================================
% Stochastic-FX variance proxy — working paper draft
% Repo: abrigo-analytics (HEAD 27fe058, plan v0.6 exit gate)
% Paper covers PRIMITIVES.md §15 open item 2 closure:
%   - PRIMITIVES.md §6 σ_T proxy and ε(σ_T) inversion as a unified mechanism
%   - Four model instantiations (deterministic + GBM + OU + Merton)
%   - cadCAD wrapping (bridge to dynamic-system simulation)
%   - AI-cost-factor model as the immediate downstream consumer
%   - Per-family canonical-pin verification (Phase A + Phase B + Phase C)
% Skill: /latex-econ-model (notation conventions + structure)
% =========================================================================
\documentclass[11pt,letterpaper]{article}

% --- Essential math packages (per latex-econ-model skill) ----------------
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{bm}
\usepackage{dsfont}

% --- Document hygiene -----------------------------------------------------
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage[margin=1in]{geometry}
\usepackage{parskip}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue!50!black,
    urlcolor=blue!50!black,
    citecolor=blue!50!black,
}
\usepackage{booktabs}
\usepackage{array}
\usepackage{graphicx}
\graphicspath{{../../notes/stochastic_fx_figures/}}

% --- Theorem environments -------------------------------------------------
\newtheorem{definition}{Definition}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{remark}{Remark}[section]
\newtheorem{assumption}{Assumption}[section]

% --- Standard econ commands -----------------------------------------------
\newcommand{\E}{\mathbb{E}}
\newcommand{\Var}{\text{Var}}
\newcommand{\Cov}{\text{Cov}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\ind}{\mathds{1}}
\newcommand{\pd}[2]{\frac{\partial #1}{\partial #2}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

% --- Project-specific commands --------------------------------------------
% σ_T realized-variance proxy (PRIMITIVES.md §6 eq. 7)
\newcommand{\sigmaT}{\sigma_{T}}
% ε inversion (PRIMITIVES.md §6 eq. 8)
\newcommand{\epsT}{\varepsilon(\sigmaT)}
% FX path numerator / denominator
\newcommand{\Xbar}{\bar{X}}
\newcommand{\Ybar}{\bar{Y}}
\newcommand{\xbar}{\bar{x}}
% (Y, M, X) triple notation
\newcommand{\Ymx}{(Y,\,M,\,X)}
% Centering projection M = I - (1/N) 1 1^T
\newcommand{\centerProj}{\mathcal{M}}
% Pin tags
\newcommand{\PinZ}[1]{\textsc{pin Z1.#1}}

\title{%
    The $\sigmaT$ Variance Proxy and Its $\varepsilon$ Inversion: \\
    A Unified Mechanism Across Deterministic and Stochastic FX Dynamics \\
    \large with a cadCAD-wrapping Bridge to the AI-Cost-Factor Downstream Consumer
}
\author{%
    Juan Manuel Serrano \\
    \texttt{juan.serranotmf@gmail.com}
}
\date{2026-05-16 \quad (working paper draft)}

\begin{document}
\maketitle

\begin{abstract}
We close \textsf{notes/PRIMITIVES.md} \S15 open item 2 (\emph{stochastic-FX
variant — Path A v3}) by verifying, across three SDE families (geometric
Brownian motion, Ornstein--Uhlenbeck, and Merton jump-diffusion), that the
realized-variance proxy
\(\sigmaT = T^{-1} \sum_{j} (X_j - \bar X_{\text{path}})^2\)
admits a single family-agnostic inversion
\(\varepsilon(\sigmaT) = \sqrt{8\,\sigmaT/\xbar^{2}}\)
that holds pointwise (\PinZ{3a}, algebraic identity) and distributionally
(\PinZ{3b} mean-only moment match against literature-derived discrete-statistic
analytic moments; \PinZ{4} Kolmogorov--Smirnov goodness-of-fit against
per-family-dispatched reference distributions). Phase B (moment) and
Phase C (KS) verdicts pass at canonical pin under each SDE family at a fixed
\(N = 1000\) Monte-Carlo budget (\PinZ{5} anti-fishing floor); audit-block
hash determinism (\PinZ{2}) is bit-exact across both intra-process and
inter-process re-runs.

The unified mechanism is wrapped, in a forthcoming integration, into a
cadCAD dynamic-agent simulation where the verified $\sigmaT$ machinery
becomes the structural M-side substrate for downstream
\(\Ymx\)-triple iterations. We position the AI-cost-factor model
(\textsf{docs/specs/2026-05-16-ai-cost-factor-model-design.md} v0.2.0,
under construction) as the immediate downstream consumer: its candidate-X
identification step is independent of the verification this paper provides,
but its eventual M-design step inherits the \(\sigmaT\)-replicating
Carr--Madan strip (\textsf{cohort\_5\_strip}, deterministic instantiation)
and the per-family stochastic verification floor.

The paper makes no causal claim about the wage-to-productive-capital
ratchet that motivates the broader Abrigo operating framework; the
contribution is strictly the verification floor on the $\sigmaT$ mechanism
that any downstream \(\Ymx\) instrument family must clear before its
M-design step is admissible.
\end{abstract}

\tableofcontents

% =========================================================================
\section{Unified framework: $\sigmaT$ as a structural primitive}
\label{sec:framework}
% =========================================================================

\subsection{Operating context and the $\Ymx$-triple unit of work}
\label{sec:framework:ymx}

The Abrigo operating framework, summarized in \texttt{CLAUDE.md}, takes as
its highest-level goal the minimization of income inequality framed in
post-Keynesian terms (distribution as institutionally determined rather
than equilibrium-given). The product family pursued by the project is
\emph{permissionless on-chain perpetual convex instruments} settled on
Panoptic, with denomination chosen per target population (Mento-native
\textsf{COPm}/\textsf{BRLm}/\textsf{KESm}/\textsf{EURm}/\textsf{USDm},
USDC, ETH, or sectoral baskets). The transmission channel from wage to
productive capital runs through a \emph{premium-funded ratchet}: a wage
earner pays a small recurring premium out of wage income, and the
instrument's accumulated convex payoff and roll yield convert over time
into productive-capital exposure that absent the instrument the wage
earner would never cross the wage--capital boundary to obtain.

The operating unit of work is the \(\Ymx\)-triple:
\begin{itemize}
    \item $Y$ — outcome variable measuring a target population's exposure
        to the wage-to-capital transition;
    \item $M$ — the Panoptic pool configuration hosting the hedge
        (underlying pair, strike/range geometry, payoff shape);
    \item $X$ — the major risk currently blocking that transition for
        the target population.
\end{itemize}

This paper does not introduce a new \(\Ymx\)-triple. It verifies a
\emph{shared substrate} that every \(\Ymx\)-triple containing a convex
on-chain hedge inherits.

\subsection{The $\sigmaT$ proxy and the $\varepsilon$ inversion}
\label{sec:framework:sigmaT}

\begin{definition}[Realized-variance proxy, $\sigmaT$]
\label{def:sigmaT}
Let $\{X_j\}_{j=0}^{n}$ denote a discrete FX-rate path observed at
$n+1$ equally spaced timepoints $t_j = j \cdot \Delta t$ on the
trading horizon $[0, T]$, with $\Delta t = T/n$ and
$\bar{X}_{\text{path}} = (n+1)^{-1}\sum_{j=0}^{n} X_j$ the path's
sample mean. The discrete realized-variance proxy is
\begin{equation}
    \sigmaT \;\coloneqq\; \frac{1}{T}\sum_{j=0}^{n}
    \bigl(X_j - \bar{X}_{\text{path}}\bigr)^{2}.
    \label{eq:sigmaT}
\end{equation}
\end{definition}

Equation~\eqref{eq:sigmaT} is the discrete-time form of
\textsf{notes/PRIMITIVES.md} \S6 eq.~(7). The normalization is by $T$
(not by $n+1$): the statistic carries units of $[X]^{2}$, and it scales
proportionally to grid density. This is intentional and load-bearing:
under any continuous-time FX-path generator, the sum-of-squared
deviations sample more variation as $n \to \infty$, and dividing by $T$
preserves comparability across grids at fixed $T$.

\begin{definition}[$\varepsilon$ inversion]
\label{def:eps}
Given a realized-variance proxy $\sigmaT$ and an FX mean $\xbar$
(typically the canonical pin $\xbar = \bar{X}/\bar{Y} = 4000$ for the
COP/USD-denominated cohort), the inversion is
\begin{equation}
    \epsT \;=\; \sqrt{\frac{8\,\sigmaT}{\xbar^{2}}}.
    \label{eq:eps}
\end{equation}
\end{definition}

Equation~\eqref{eq:eps} is \textsf{PRIMITIVES.md} \S6 eq.~(8). It is
purely algebraic: no SDE-family parameter enters; the relation is
family-agnostic. The square-root form ensures (\PinZ{3a} trivial-degenerate
limit) that
\(
    \lim_{\sigmaT \to 0^{+}} \epsT \;=\; 0,
\)
which captures the geometric fact that a flat (deterministic) path has
zero implied volatility per unit FX mean.

\subsection{Verification structure: three phases, six pins}
\label{sec:framework:phases}

The verification this paper reports has three phases, each anchored to a
distinct numbered pin in
\textsf{docs/specs/2026-05-11-stochastic-fx-variant-design.md} v0.5:

\begin{description}
    \item[Phase A — algebraic identity (\PinZ{3a}).] For each path's
        $\sigmaT^{(n)}$, the pointwise identity
        $\bigl(\varepsilon(\sigmaT^{(n)})\bigr)^{2} = 8\sigmaT^{(n)} / \xbar^{2}$
        holds to float64 precision (\textsc{numerical\_identity\_tol} $= 10^{-6}$).
        Family-agnostic; reduces to a re-evaluation of~\eqref{eq:eps}.

    \item[Phase B — mean-only moment match (\PinZ{3b}, spec v0.5 amendment).]
        Empirical mean of $\sigmaT^{(n)}$ across the $N$-path ensemble
        matches the hand-derived analytic $\E[\sigmaT]$ for the
        DISCRETE statistic~\eqref{eq:sigmaT} at the canonical grid,
        within $\textsc{moment\_rel\_tol} = 0.05$. The discrete-statistic
        analytic uses the centering-projection identity
        \(
            \E[\sigmaT \cdot T] = \mathrm{tr}\bigl(\centerProj\,\Sigma\bigr)
                                  + \mu^{\top} \centerProj\, \mu
        \)
        where $\centerProj = I_{n+1} - (n+1)^{-1}\mathbf{1}\mathbf{1}^{\top}$
        and $\Sigma$ is the SDE-family-specific auto-covariance kernel
        on the canonical grid. Variance is NOT in this gate (see
        Remark~\ref{rem:mean-only}).

    \item[Phase C — Kolmogorov--Smirnov goodness-of-fit (\PinZ{4},
        spec v0.5 amendment).] Per-family dispatch on the reference
        distribution shape (not on the threshold):
        \begin{itemize}
            \item GBM: method-of-moments lognormal reference fit to the
                analytic $(\E, \Var)$, tested via
                \texttt{scipy.stats.ks\_1samp}.
            \item OU: method-of-moments gamma reference fit to the
                analytic $(\E, \Var)$, tested via \texttt{ks\_1samp}.
            \item Merton: empirical-CDF reference via a high-$N$
                ($N_{\textsc{ref}} = 100\,000$) reference run from the
                same SDE, tested via \texttt{ks\_2samp}. This avoids the
                shape mismatch between Merton's Poisson-mixture-of-lognormals
                geometry and the 2-parameter lognormal alternative.
        \end{itemize}
        Single threshold across all three families:
        $\textsc{ks\_pvalue\_floor} = 0.01$.
\end{description}

\begin{remark}[Why mean-only Phase B; what variance discloses]
\label{rem:mean-only}
At the \PinZ{5} anti-fishing path-count floor $N = 1000$, the standard
error of the sample-variance estimator on $\sigmaT$ is intrinsically
$\sqrt{(\kappa_{\text{eff}} - 1)/(N-1)}$ where $\kappa_{\text{eff}}$
ranges 4--35 across the three families (driven by jump kurtosis for
Merton). This produces an 8\%--30\% Monte-Carlo noise floor on the
variance estimator — strictly incompatible with
$\textsc{moment\_rel\_tol} = 0.05$. The spec v0.4 attempt to gate
variance against the same 5\% tolerance was mathematically
unsatisfiable; v0.5 \S11.6 dropped variance from the Phase B gate.
Variance relative error is still computed and emitted on the
\textsf{InversionVerdict} as an audit-trail observation (it is not
zero); the full-distribution match is preserved at Phase C, where the
KS test against a moment-matched reference constrains the joint shape
(mean, variance, skewness, kurtosis) via the empirical CDF.
\end{remark}

\subsection{Pin coverage and anti-fishing posture}
\label{sec:framework:pins}

The six pins anchoring this paper are:

\begin{center}
\begin{tabular}{@{}lp{0.7\textwidth}@{}}
\toprule
\textbf{Pin} & \textbf{Statement} \\
\midrule
\PinZ{1} & SDE parameters pre-pinned ex-ante per family (Definition~\ref{def:canonical-pins} below). \\
\PinZ{2} & Path-ensemble determinism: identical
            $(\text{params}, \text{rng\_seed}, n_{\text{paths}})$
            triples yield bit-exact paths and bit-exact audit\_block. \\
\PinZ{3a} & Algebraic identity \(\bigl(\epsT\bigr)^{2} = 8\sigmaT/\xbar^{2}\)
            to float64 precision. \\
\PinZ{3b} & Mean-only moment match against discrete-statistic analytic
            $\E[\sigmaT]$ within 5\%. \\
\PinZ{4} & Per-family Phase C KS p-value $\geq 0.01$. \\
\PinZ{5} & Anti-fishing routing: failures route to \textsc{corrections-$\alpha$}
            and scoped two-reviewer re-review; NEVER silent parameter
            re-tuning; NEVER ``increase $N$ until passing''; \(N=1000\)
            is an invariant, not a parameter to grow. \\
\PinZ{6} & Strip preservation: the
            \textsf{cohort\_5\_strip}/\textsf{IronCondor\_strip.json}
            audit\_block remains byte-equal across the verification
            plan's execution window. \\
\bottomrule
\end{tabular}
\end{center}

\PinZ{5} carries forward the broader project's anti-fishing invariants
($N_{\min} = 75$, $\textsc{power}_{\min} = 0.80$,
$\textsc{mdes\_sd} = 0.40$ SD-units of $Y$, pre-pinned sign expectation
and lag structure). For this paper specifically, three concrete
anti-fishing prohibitions hold across all four model instantiations:
(i)~no silent threshold relaxation, (ii)~no seed re-roll on failure,
(iii)~no $N$-floor lowering. The four spec amendments
(\S11.6, \S11.7 of the parent spec; \S16.3, \S16.4, \S16.5, \S16.7 of
the plan) are each subjected to scoped two-reviewer re-review per
\textsf{master-spec} \S6.4, and each preserves these three prohibitions.

\subsection{Downstream consumer: the AI-cost-factor model}
\label{sec:framework:downstream}

The verification floor established by this paper is the immediate
prerequisite for an \(\Ymx\)-triple targeting the COP-denominated AI-tooling
cost burden borne by non-subscribed LATAM developers (the AI-cost-factor
model under construction, spec v0.2.0; see
\S\ref{sec:cadcad:cost-factor}). That model's X-channel identification
step (R5 descriptive risk quantification + R4-Stage-3 vol-clustering
regression) operates on the empirical cost path, NOT on $\sigmaT$
directly. But should X-identification yield a candidate FX-vol risk
that admits Panoptic-eligible M-design, the M-side instrument is a
$\sigmaT$-replicating strip and inherits the per-family stochastic
verification floor of this paper.

This positioning is asymmetric on purpose: the $\sigmaT$ machinery is
\emph{substrate} for downstream \(\Ymx\)-triple iterations, not the
final product. The contribution of this paper is the substrate
verification; downstream consumers carry their own identification
burdens.

% =========================================================================
\section{Deterministic instantiation: Carr--Madan replication}
\label{sec:deterministic}
% =========================================================================

\paragraph{Section charge for content fill.}
This section instantiates the framework's first concrete model: the
deterministic FX-path generator \textsf{PRIMITIVES.md} \S5 eq.~(6)
$f(t) = E^{-2} |\cos(E t)|$ and the corresponding $\sigmaT$-replicating
Carr--Madan 12-leg IronCondor strip committed at
\textsf{simulations/saas\_builder/cohort\_5\_strip/} (audit\_block
\texttt{94150326332b90e5\ldots aaf329}, \PinZ{6}). Cite
\textsf{notes/STAGE\_2\_RESULTS.md} R1 (\(\sigma_0\) anchor) for the
deterministic-limit reduction.

% =========================================================================
\section{Stochastic instantiation I: geometric Brownian motion}
\label{sec:gbm}
% =========================================================================

\paragraph{Section charge for content fill.}
Section presents the GBM SDE
$dX_t = \mu X_t \,dt + \sigma X_t \,dW_t$
with the Euler--Maruyama log-scale discretization (Itô correction
$-\sigma^{2}/2$); the discrete-statistic analytic $\E[\sigmaT]$ via the
centering-projection identity against the GBM auto-covariance kernel
$\Cov(X_j, X_k) = x_{0}^{2}\bigl(\exp(\sigma^{2} \min(t_j, t_k)) - 1\bigr)$
at $\mu = 0$ canonical; verification: Phase A residual $\sim 7 \times 10^{-15}$,
Phase B mean rel-err $1.12\%$, Phase C lognormal-MoM KS p-value $0.123$.
LaTeX fragment in \textsf{notes/stochastic\_fx\_tex/sigma\_t\_moments\_gbm.tex}.
Cite Hull (Options, Futures, and Other Derivatives, \S15).

% =========================================================================
\section{Stochastic instantiation II: Ornstein--Uhlenbeck}
\label{sec:ou}
% =========================================================================

\paragraph{Section charge for content fill.}
Section presents the OU SDE
$dX_t = \theta(\bar\mu - X_t)\,dt + \sigma\, dW_t$
with the exact Vasicek transition (NOT Euler); stationary auto-covariance
$\Cov(X_j, X_k) = (\sigma^{2}/(2\theta))\,\exp(-\theta|t_k - t_j|)
\bigl(1 - \exp(-2\theta \min(t_j, t_k))\bigr)$ at $x_0 = \bar\mu$
canonical (so the mean term vanishes); verification: Phase A residual
$\sim 4 \times 10^{-15}$, Phase B mean rel-err $0.17\%$, Phase C
gamma-MoM KS p-value $0.303$.
LaTeX fragment in \textsf{notes/stochastic\_fx\_tex/sigma\_t\_moments\_ou.tex}.
Cite Karatzas \& Shreve (Brownian Motion and Stochastic Calculus, \S5.6).

% =========================================================================
\section{Stochastic instantiation III: Merton jump-diffusion}
\label{sec:merton}
% =========================================================================

\paragraph{Section charge for content fill.}
Section presents the Merton jump-diffusion SDE
$dX_t = \mu X_t\,dt + \sigma X_t\,dW_t + X_{t^-}(J - 1)\,dN_t$
where $N_t$ is a Poisson process of intensity $\lambda$ and $J$ is
lognormal; Euler--Maruyama log-scale diffusion augmented by compound-Poisson
log-jumps; conditional-Gaussian aggregation via
$\text{aggregated\_log\_jump} = N \mu_{J} + \sqrt{N}\, \sigma_{J}\, Z'$;
discrete-statistic analytic $\E[\sigmaT]$ with non-constant drift
$\mu_{j} = x_{0}\,\phi_{1}^{j}$ where
$\log\phi_{1} = \lambda\Delta t\bigl(\exp(\sigma_{J}^{2}/2) - 1\bigr)$;
verification: Phase A residual $\sim 2 \times 10^{-13}$, Phase B mean
rel-err $0.70\%$, Phase C empirical-CDF KS-2-samp p-value $0.116$
(at hardware-constrained $N_{\textsc{ref}} = 5000$; production
$N_{\textsc{ref}} = 100\,000$).
LaTeX fragment in \textsf{notes/stochastic\_fx\_tex/sigma\_t\_moments\_merton.tex}.
Cite Andersen \& Piterbarg (Interest Rate Modeling Vol.~I, \S2.7).

% =========================================================================
\section{cadCAD wrapping: bridge to dynamic-system simulation}
\label{sec:cadcad}
% =========================================================================

\subsection{Motivation and scope}
\label{sec:cadcad:motivation}

\paragraph{Section charge for content fill.}
Subsection explains why a cadCAD wrapping is necessary even after the
$\sigmaT$ machinery verifies under three SDE families: the verification
floor establishes that the inversion mechanism holds, but downstream
\(\Ymx\)-triple iterations need to simulate the interaction of multiple
state variables (FX path, cost path, replicating-portfolio position,
hedge accumulation) over time. The cadCAD framework
(\textsf{cadCAD-org/cadCAD}) provides a dynamic-agent simulation engine
with state variables, policy functions, and partial state update blocks
(PSUBs) suitable for this composition.

\subsection{State variables and PSUB design}
\label{sec:cadcad:psub}

\paragraph{Section charge for content fill.}
Subsection introduces the cadCAD state-variable signature: at each time
step $t$, the simulation tracks the FX path $X_t$ (sampled from one of
the three SDE families plus the deterministic baseline), the cost
trajectory $C_t$ (for the AI-cost-factor model, downstream consumer),
the replicating-portfolio's $\sigmaT$ accumulation, the wage-funded
premium contribution rate, and the convex-payoff-to-productive-capital
ratchet variable. PSUBs:
(i)~FX-path-advance (SDE-family-dispatched);
(ii)~$\sigmaT$ accumulator update;
(iii)~premium-payment-and-position-update;
(iv)~ratchet-conversion at threshold crossings.

\subsection{The AI-cost-factor model as downstream consumer}
\label{sec:cadcad:cost-factor}

The AI-cost-factor model specified at
\textsf{docs/specs/2026-05-16-ai-cost-factor-model-design.md} v0.2.0
(currently under construction; v0.1.0 rejected by all three reviewers
with five \textsc{block}s on identification and unitization grounds)
targets the COP-denominated AI-tooling cost burden borne by
non-subscribed LATAM developers. The cost-factor's identification stage
(\S2.1 R5 descriptive risk quantification + \S2.2 R4-Stage-3 vol-clustering
regression on log-COP-cost-returns) is independent of the
verification this paper provides; PASS on the cost-factor's
identification opens an M-design step that consumes the verified
$\sigmaT$-replicating mechanism.

The coupling in the cadCAD simulation is asymmetric:
\begin{enumerate}
    \item The FX-path PSUB (one of the four SDE families) produces $X_t$.
    \item The cost-trajectory PSUB consumes $X_t$ and emits $C_t$ as a
        function of (token usage demand, API rate-card, $X_t$).
    \item The $\sigmaT$ accumulator (PRIMITIVES.md \S6 eq.~(7) discretely)
        consumes the FX-path window and emits the realized variance.
    \item The replicating-portfolio PSUB (Carr--Madan 12-leg IronCondor
        instantiation) consumes $\sigmaT$ and the inversion
        $\epsT$ to construct the hedge position.
    \item The wage-funded premium PSUB drains a fixed share of the
        cohort's wage income each step, recapitalizing the position.
    \item The ratchet variable accumulates the convex payoff and crosses
        the productive-capital-equivalence threshold per the cohort's
        denomination.
\end{enumerate}

This paper does not implement the cadCAD wrapping; the implementation
work is the immediate next deliverable, specified to follow doc-first
sequencing from this paper's framework. The cost-factor model's R5/R4-S3
results, when they land, will be reported in a companion paper.

\subsection{What the cadCAD wrapping does NOT do}
\label{sec:cadcad:nondoes}

\paragraph{Section charge for content fill.}
Subsection lists the explicit exclusions: cadCAD wrapping does NOT
re-validate the $\sigmaT$ machinery (that's this paper's job); does NOT
estimate the cost-factor model's identification parameters (that's the
cost-factor spec's job); does NOT cross the wage-to-capital boundary
empirically (the ratchet variable's threshold crossings are simulated
under the cohort's assumed denomination, not empirically observed).

% =========================================================================
\section{Results: canonical-pin verification per family}
\label{sec:results}
% =========================================================================

\paragraph{Section charge for content fill.}
Section reports the canonical-pin verification verdict per family,
sourced from \textsf{notes/STOCHASTIC\_FX\_RESULTS.md} (the
authoritative Stage-3 open-item 2 closure document). Cross-family
summary table; per-family $(\sigma, \theta, \lambda, \sigma_J)$ parameter
pins; per-family $(\text{Phase A residual}, \text{Phase B mean rel-err}, \text{Phase B var rel-err (audit-trail)}, \text{Phase C KS p-value}, \text{audit\_block})$
tuples. Inline cross-check via \textsf{scratch/2026-05-11-stochastic-fx-execution/driver.py}
re-runnable from a fresh checkout (\PinZ{2} cross-process determinism).
Per-family interpretation paragraphs: GBM (lognormal $\sigmaT$ shape
absorbs both first and second moments); OU (Gaussian quadratic form
yields exact $\E[\sigmaT] = \sigma^{2}/(2\theta)$ at $x_0 = \bar\mu$;
gamma reference fits cleanly); Merton (mixture geometry requires the
empirical-CDF Phase C dispatch per spec v0.5 \S11.7).

% =========================================================================
\section{Discussion: what the verification floor unlocks}
\label{sec:discussion}
% =========================================================================

\paragraph{Section charge for content fill.}
Section synthesizes the four model-instantiation results into a single
statement of what the verification floor unlocks for downstream
\(\Ymx\)-triple iterations. Three honest limitations: (i)~mean-only
Phase B at $N = 1000$ has a 0\%--30\% Type-I rate per family
(spec v0.5 \S11.6 disclosure); (ii)~the Merton Phase C reference at
hardware-constrained $N_{\textsc{ref}} = 5000$ matches the production
$N_{\textsc{ref}} = 100\,000$ in the same KS p-value ballpark per the
Wave-2 robustness probe (KS p $\approx 0.12$--$0.14$), but production
hardware $\geq 8$ GB free is the binding constraint for the canonical
reference; (iii)~the deterministic eq.~(6) generator produces synchronized
paths whose ergodic content is artificial, motivating the stochastic
extension this paper closes. Forward statement: AI-cost-factor model
(downstream consumer) inherits this floor; cadCAD wrapping is the next
deliverable; an instrument-deployment study under live Panoptic
liquidity is gated on M-design step success in a future \(\Ymx\)-iteration
not yet specified.

% =========================================================================
\appendix
% =========================================================================

\section{Per-family analytic-moment derivations}
\label{app:moments}

\paragraph{Appendix charge for content fill.}
Per-family derivation of $\E[\sigmaT]$ for the DISCRETE statistic
\eqref{eq:sigmaT} via the projection identity
$\E[\sigmaT \cdot T] = \mathrm{tr}(\centerProj\,\Sigma) + \mu^{\top}\centerProj\,\mu$.
Sub-appendices A.1 GBM, A.2 OU, A.3 Merton. Each cites the textbook
auto-covariance kernel + transcribes the centering-projection
computation. The scratch derivation script
\textsf{scratch/2026-05-13-task-4.2-discrete-moments/derivation.py} is
the load-bearing artefact.

\section{Anti-fishing audit trail across spec amendments}
\label{app:antifishing}

\paragraph{Appendix charge for content fill.}
Appendix walks the four spec amendments (\S11.6, \S11.7 of the parent
spec; \S16.3, \S16.4, \S16.5, \S16.7 of the plan) and confirms each
preserves the three concrete anti-fishing prohibitions (no silent
threshold relaxation, no seed re-roll, no $N$-floor lowering). Each
amendment's Wave-1 RC + MQ Delphi-independent re-review verdict files
are cited; Wave-2 RC + MQ at plan-exit are cited.

\section{cadCAD-implementation spec (forward-looking)}
\label{app:cadcad-spec}

\paragraph{Appendix charge for content fill.}
Appendix sketches the cadCAD implementation spec that this paper
nominates for the immediate next deliverable: directory structure
(\textsf{simulations/cadcad/}), state-variable declarations, PSUB
implementations per the six PSUB roles itemized in
\S\ref{sec:cadcad:psub}, integration tests crossing the four FX
families and the AI-cost-factor cost trajectory, anti-fishing
invariants carried forward.

% =========================================================================
% References
% =========================================================================
\bibliographystyle{plain}
% TODO: produce notes/papers-bib/stochastic-fx-paper.bib with Hull,
%       Karatzas-Shreve, Andersen-Piterbarg, Carr-Madan, Merton (1976),
%       Vasicek (1977), Itô, cadCAD whitepaper, Panoptic whitepaper,
%       post-Keynesian distribution literature.
\begin{thebibliography}{10}
\bibitem{hull-2017}
    Hull, John C. (2017). \emph{Options, Futures, and Other Derivatives},
    10th ed. Pearson. \S15.

\bibitem{karatzas-shreve-1991}
    Karatzas, Ioannis \& Steven E. Shreve (1991).
    \emph{Brownian Motion and Stochastic Calculus},
    2nd ed. Springer. \S5.6 (Ornstein--Uhlenbeck process).

\bibitem{andersen-piterbarg-2010}
    Andersen, Leif B. G. \& Vladimir V. Piterbarg (2010).
    \emph{Interest Rate Modeling, Volume I: Foundations and Vanilla Models}.
    Atlantic Financial Press. \S2.7 (jump-diffusion moments).

\bibitem{carr-madan-2001}
    Carr, Peter \& Dilip Madan (2001).
    Optimal Positioning in Derivative Securities.
    \emph{Quantitative Finance} 1(1), 19--37.

\bibitem{merton-1976}
    Merton, Robert C. (1976).
    Option Pricing When Underlying Stock Returns Are Discontinuous.
    \emph{Journal of Financial Economics} 3, 125--144.

\bibitem{vasicek-1977}
    Vasicek, Oldrich (1977).
    An Equilibrium Characterization of the Term Structure.
    \emph{Journal of Financial Economics} 5(2), 177--188.

\bibitem{cadcad}
    cadCAD: complex Adaptive Dynamics Computer-Aided Design.
    \url{https://github.com/cadCAD-org/cadCAD}.

\bibitem{panoptic-2023}
    Panoptic: Perpetual Options on Uniswap.
    Technical whitepaper. 2023.

\bibitem{abrigo-stochastic-fx-spec}
    Abrigo Analytics (2026). Stochastic-FX variant design spec, v0.5.
    \texttt{docs/specs/2026-05-11-stochastic-fx-variant-design.md},
    in \texttt{abrigo-analytics} repository.

\bibitem{abrigo-stochastic-fx-plan}
    Abrigo Analytics (2026). Stochastic-FX variant implementation plan, v0.6.
    \texttt{docs/plans/2026-05-11-stochastic-fx-variant.md},
    in \texttt{abrigo-analytics} repository.

\bibitem{abrigo-ai-cost-factor}
    Abrigo Analytics (2026). AI-Cost Factor Model design spec, v0.2.0
    (under construction).
    \texttt{docs/specs/2026-05-16-ai-cost-factor-model-design.md},
    in \texttt{abrigo-analytics} repository.

\end{thebibliography}

\end{document}
